### 2024-10-31 main-disk-model.tex

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{algorithmicx}
%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{numprint}
\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{balance}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\include{macros}

\newcommand{\algo}[1]{\textsc{#1}}
\newcommand{\bottomlevel}[1]{\underline{l}_{#1}} % underline short italic
\newcommand{\criticalpath}{\mathcal{P}}
\newcommand{\parents}[1]{\,\Pi_{#1}}
\newcommand{\children}[1]{\,C_{#1}}
\newcommand{\cluster}{\,\mathcal{S}}

\newcommand{\heft}{\algo{HEFT}\xspace}
\newcommand{\heftmm}{\algo{HEFTM-MM}\xspace}
\newcommand{\heftbl}{\algo{HEFTM-BL}\xspace}
\newcommand{\heftblc}{\algo{HEFTM-BLC}\xspace}

\newcommand{\MM}{M}
\newcommand{\MC}{MC}
\newcommand{\rt}{rt}
\newcommand{\curM}{curM}
\newcommand{\curC}{curC}
\newcommand{\PD}{PD}

\newcommand{\bw}{bw}
\newcommand{\br}{br}

\newcommand{\skug}[1]{{\color{blue}[SK: #1]}}
\newcommand{\hmey}[1]{{\color{red}[HM: #1]}}
\newcommand{\AB}[1]{{\color{purple}[AB: #1]}}
\newcommand{\willchange}[1]{{\color{orange}[AB: #1]}}

\renewcommand{\iec}{i.e., }
\begin{document}

    \title{Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\
%{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
%should not be used}
    % \thanks{Identify applicable funding agency here. If none, delete this.}
    }

%\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID} }

    \maketitle

    \begin{abstract}

        %Scheduling scientific workflows is important. (\skug{can't imagine a good first sentence})
        Scientific workflows are often represented as directed acyclic graphs (DAGs),
        where vertices correspond to tasks and edges represent the dependencies between them.
        Typically, each task requires a
        certain amount of memory to be executed and needs to communicate data to its successor tasks.
        The goal is generally to execute the workflow  as fast as possible (i.e., to minimize its makespan),
        while satisfying the memory constraints.
        Hence, we investigate the memory-aware scheduling of DAG-shaped workflows on
        heterogeneous platforms, where each processor can have a different speed and a different memory size.
        We propose a variant of HEFT (Heterogeneous Earliest Finish Time) that (in contrast to the original) accounts for memory and
        includes eviction strategies for cases when it might be beneficial to remove some data from memory
	in order to have enough memory to execute other tasks.
%
	Furthermore, while HEFT assumes perfect knowledge of the execution time and memory usage
	of each task, the actual values might differ upon execution. Thus, we propose an adaptive
	scheduling strategy, where a schedule is recomputed when there has been a significant variation in terms
	of execution time or memory.
%
	The scheduler has been closely integrated with a runtime system, allowing us to perform a thorough
	experimental evaluation on real-world workflows. The runtime system warns the scheduler when
	the task parameters have changed, and a schedule can be recomputed on the fly. The memory-aware
	strategy allows us to schedule task graphs that would run out of memory with a state-of-the-art
	scheduler, and the adaptive setting allows us to significantly reduce the makespan.

%                when tentatively assigning tasks, in order to
%        Its first step is to compute the weights of the tasks.
%        We suggest three variants: bottom levels as weights, bottom levels with impact of incoming edge weight,
%        and weights along the optimal memory traversal.
%        In the second step, we try assigning each task to each processors and execute the assignment that
%        is feasible with regard to memory size and gives the earliest finishing time to the task.
%        Sometimes, data corresponding to edge weights that is stored in the memory needs to be evicted in order to
%        assign a task to a processor.
%        We suggest two eviction strategies - largest files first and smallest files first.

%        Our experimental evaluation on real-world workflows and simulated (\skug{generated? They are generated from real-world wfs})
%        ones with real task and edge weights with up to 30,000 tasks shows that
%        respecting memory constraints only costs $11\%$ of runtime in comparison to a non memory-aware baseline.
%        Calculating task weights with the impact of memory gives a $x\%$ better makespans on a normal and $y\%$ better makespans
%        on small one.
%        Calculating task weights along the optimal memory traversal gives on average $z\%$ worse makespans, but improves
%        average memory utlization by $t\%$.

    \end{abstract}

%    \begin{IEEEkeywords}
%        DAG, Heterogeneous platform, Adaptive scheduling, Memory constraint.
%    \end{IEEEkeywords}

\section{Introduction} %: \skug{Full: 0, Polished: 0}}

%    \skug{Fullness score (how much text is available): 0-none to 5 - everything.
%
%    Polishedness score (how well-written is the chapter): 0 messy - 5 very clean.
%    }
%    TODO: Insert abstract

    %%% CONTEXT %%%
    The analysis of massive datasets, originating from fields such as genomics, 
    remote sensing, or biomedical imaging -- to name just a few -- has become ubiquitous in science;
    this often takes the form of workflows, \iec separate software components chained together
    in some kind of complex pipeline~\cite{DBLP:journals/dbsk/LeserHDEGHKKKKK21}.
    These workflows are usually represented as directed acyclic graphs (DAGs).
    The DAG vertices represent the software components (or, more generally, the workflow \emph{tasks}),
    while the edges model I/O dependencies between the tasks~\cite{adhikari2019survey,liu2018survey}.
    Large workflows with resource-intensive tasks can easily exceed the capabilities of a 
    single computer and are therefore executed on a parallel or distributed platform.
    An efficient execution of the workflows on such platforms requires mapping tasks
    to specific processors; to increase utilization by reusing finished processors,
    one also needs a task schedule (\iec a valid execution order that respects the dependencies)
    and possibly also starting times for the tasks.

    %%% MOTIVATION %%%
    Modern computing platforms are often heterogeneous, meaning they feature varying CPU speeds
    and memory sizes. In general, having different memory sizes per CPUs makes it more challenging to compute
    a schedule that respects all memory constraints -- meaning that no task is executed on a 
    processor with less memory than needed for the task. This is, however, very important to
    avoid (possibly expensive) runtime failures and to provide a satisfactory user experience.
    Hence, building on previous related %\hmey{You may want to add other refs not from us} 
    work~\cite{gou2020partitioning,He21,DBLP:conf/icpp/KulaginaMB24}, we consider a scheduling problem 
    formulation that takes memory sizes as explicit constraints into account. Its objective is
    the very common \emph{makespan}~\cite{liu2018survey}, 
    which acts as proxy for the total execution time of a workflow.
However, to the best of our knowledge,
    the only memory-aware heuristics that would account for memory constraints are  partitioning
    the DAG and not reusing processors once they have processed a part of the graph, leading
    to high values of makespan compared to a finer grain solution that reuses processors.

    While previous work with memory constraints has been focusing on partitioning the graph, 
    and not reusing processors during execution, 
    a seminal list scheduling heuristic for workflows on heterogeneous platforms, without accounting
    for the memory constraint, is HEFT 
    (heterogeneous earliest finish time)~\cite{topcuoglu2002performance}.
    It has two phases: (i) each task is assigned a priority; and (ii) the tasks in a priority-ordered list are assigned
    to processors, where the ``ready'' task with the highest priority is scheduled next on the processor
    where it would complete its execution first. 
    HEFT has been extended (e.g., by Shi and Dongarra~\cite{SHI2006665}) and adjusted 
    for a variety of different scheduling problem formulations. 
    Yet, none of them adhere to memory constraints as we propose -- see discussion of related work
    in Section~\ref{sec:related-work}. 
%    (\skug{check in related work if true!}).
%        \skug{Note: 2 papers that deal with memory sizes, but model is very different!}

    Another limitation in practice of HEFT (and many other scheduling strategies) is their 
    assumption that the task running times provided to them are accurate. In practice, this is 
    not the case and deviations from user estimates or historical measurements are 
    very common~\cite{hirales2012multiple}. As a consequence, one should adapt the schedule when \emph{major}
    deviations occur. However, the original list-based schedulers, such as HEFT, are only defined
    in a static setting with accurate task parameters. 
%    
%    List-based schedulers such as HEFT are, however, not designed for 
%    such an adaptation~\cite{TODO}.\hmey{Svetlana, Anne: is this a fair statement? Please add ref (if any)}
%    \AB{Actually, we adapt HEFT as well, I'll reformulate...}
%%    and would compute a completely new schedule from scratch.

    %%% CONTRIBUTION %%%
    %\paragraph*{Contribution} 
    The main contributions of this paper are both algorithmic and experimental:
    \begin{itemize}
\item We formalize the problem with memory constraints, where communication buffers
are used to evict data from memory if it will be later used by another processor. 
\item We design three HEFT-based heuristics
    that adhere to memory size constraints: \heftbl, \heftblc, and \heftmm
    (M behind HEFT for \underline{m}emory, BL for \underline{b}ottom \underline{l}evel,
    BLC for \underline{b}ottom \underline{l}evel with \underline{c}ommunication, 
    and MM for \underline{m}inimum \underline{m}emory traversal).
    The difference between the new heuristics is the way they prioritize tasks (for processor assignment).

    \item We implement a runtime system able to provide some feedback to the scheduler
    when task requirements (in terms of execution time and memory) differ from the initial predictions, 
    and we recompute a schedule, based on the reported deviations.

    \item We perform extensive simulations, first in the static case by comparing the schedules produced 
    by these heuristics with the classical HEFT as baseline, which however does not take memory sizes into account; 
    while HEFT returns invalid schedules that exceed the processor memories and cannot execute correctly,
    the new heuristics are able to successfully schedule large workflows, with reasonable makespans.

    \item In the dynamic setting, we use a runtime system that allows us to simulate workflow executions,
    introducing deviations in running times and memory requirements of tasks that are communicated
    back to the scheduler; the scheduler can then recompute a schedule. Without these recomputations,
    most schedules become invalid after deviations, since the memory constraint is exceeded 
    for most workflows, hence demonstrating the necessity of a dynamic adjustment of the schedule.

%    \hmey{Need to clarify: why is this SotA?}

  %  \begin{itemize}
%      \item Static: we find that our heuristics are able to schedule all workflows correctly, and produce makespans similar to the baseline.
%      \item Adaptive: runtime system built, simulates workflow executions and deviations in running times and mem requirements of tasks
%      \item Answering requests of the runtime system for adaptation, the scheduler computes an improved schedule based on the reported deviations.
    \end{itemize}

We first review related work in Section~\ref{sec:related-work}. Then, we formalize the model in Section~\ref{sec:model}
and the algorithms in Section~\ref{sec:heuristics}. The adaptation of the heuristics in a dynamic setting is discussed in Section~\ref{sec:dyn}, and experiment results are presented in Section~\ref{sec:expe}. Finally, we conclude
and provide future working directions in Section~\ref{sec:conc}.

\section{Related work} %: \skug{Full: 5, Polished: 4}}
    \label{sec:related-work}

   First, we focus on HEFT-like scheduling heuristics from the literature  that do not necessarily  
   consider memory constraints. Then, we discuss memory-aware scheduling algorithms. 
   Finally, we move to related work on dynamic or adaptive algorithms. 
 %   \hmey{Question: Does it make sense to place related work before the model (which is more common)?}

%    We discuss relevant scheduling approaches that reuse processors or respect the memory requirements of the processors.
%
%    \subsection{Early list schedulers with unlimited processors}
%    An entire cluster of works on list schedulers has been carried out as early as the 90s.
%    They all assume a DAG-shaped workflow with makespan weights on tasks, and an unlimited amount of homogeneous processors
%    with the speed of 1~\cite{benoit2013survey}.\hmey{needs refs or at least a survey pointer}
%    \skug{actually, Anne cited them, so citing her :-)}
%
%    The \textit{task duplication}-based approaches exploit that sometimes running a task twice on different machines can
%    help reduce the makespan by saving communication costs.
%    The two categories are scheduling with partial duplication~(SPD), and with full duplication~(SFD).
%    For a join task (a task whose incoming degree is larger than its outgoing degree), SPD finds a critical immediate
%    parent (the one that gives the largest start time to the join task) and duplicates only it.
%    SFD duplicates all parents of a join node.
%    The algorithm by~\cite{dfrn1997} duplicates first (creates copies of all parent tasks) and then reduces (removes) the ones that can be removed without harming the makespan.
%    The critical path fast duplication algorithm CPFD~\cite{5727760} classifies tasks into three categories: critical
%    path task, in-branch task, or out-branch task.
%    It schedules critical path tasks first, then in-branch tasks.
%    \hmey{I'm missing the connection to our paper. Or vice versa, how is our contribution
%    connected to these works? (Einordnung in den Forschungskontext) For example, do we do similar things? Are there interesting limitations we overcome?}
%
%    Linear clustering~\cite{KWOK1999381} acts on critical paths in the workflow.
%    It assigns the current critical path to one processor, removes all these tasks from the workflow, recomputes the critical
%    path and repeats the procedure.
%    Heaviest node first~\cite{SHIRAZI1990222} assigns the tasks level by level;
%    in each level, it schedules the one with largest computation time first.

    \subsection{Static list schedulers, especially HEFT-based algorithms}

    Introduced in 2002, HEFT~\cite{topcuoglu2002performance} is a list-based heuristic.
    It and all its successors consist of two phases: task prioritization/ordering and task assignment.
    In the first phase, the algorithms compute bottom levels of the tasks based on some priorities (create the list),
    and then schedule tasks in the order of these priorities.
    The modifications of HEFT revolve around the way the priorities of the tasks are computed and the logic of the processor assignment.
    All such algorithms assume a heterogeneous execution environment.

    Hence, during the task prioritization phase in~\cite{sulaiman2021hybrid},  the standard deviation of the computation cost
    (between processors) is computed, and added to the mean value to account for the differences between processor speeds.
    In the processor choice phase, the entry task and the longest predecessor tasks are duplicated 
    during idle times on the processor.

%    Ref.~\cite{alebrahim2017task} computes the bottom level based on the difference of execution times on
%    the fastest and the slowest processors, divided by the speed ratio of these two processors.
%    When doing processor selection, the authors differentiate between the lowest execution time and earliest finishing time.
%    They choose the processor with the lowest execution time and cross over to other processors sometimes.
%    They build upon~\cite{shetti2013optimization}.\hmey{Last sentence ``hangs in the air''. Either drop or connect it properly.}

    PEFT (Predict earliest finish time)~\cite{arabnejad2014list} is a HEFT variant that computes an Optimistic
    Cost Table (OCT).
    The OCT is computed per task-processor pair and stores the longest shortest path from this task to the target task if this
    processor is chosen for this task.
    Ranking is based on OCT values.
    The processor choice stage minimizes the optimistic EFT, which is EFT plus the longest path to the exit node for each task.

    The HSIP (Heterogeneous Scheduling with Improved task Priorities)~\cite{wang2016hsip} has an improved first step in
    comparison to HEFT.
    It combines the standard deviation with the communication cost weight on the tasks.
    In the second stage, the algorithm duplicates the entry task if there is a need for it.

    The TSHCS (Task Scheduling for Heterogeneous Computing Systems) algorithm~\cite{alebrahim2017task} improves on HEFT
    by adding randomized decisions to the second phase.
    The decision is whether the task be assigned to the processor with the lowest execution time or to the processor that
    produces the lowest finish time.

    The SDC algorithm~\cite{SHI2006665} considers the percentage of feasible processors in addition to task’s
    average execution cost in its weight.
    The selected task is then assigned to a processor which minimizes its Adjusted Earliest Finish Time (AEFT) that
    additionally notes how large the communication between current node and its children will be on the
    average provided that it is scheduled on the current processor.

    HEFT  can also be adapted in cloud-oriented environments~\cite{samadi2018eheft} and even combined with reinforcement learning techniques~\cite{yano2022cqga}.

    \subsection{Memory-aware scheduling algorithms}
    Respecting processor memories adds a constraint to a scheduling problem.
    Therefore, only specifically memory-targeted algorithms address this issue.
    Moreover, the way processor memories are represented in the model has a decisive impact on the way the constraint
    is formulated and addressed in the algorithm.
%
    Different models of memory available on processors and memory requirements of tasks have been presented.

    Marchal~et~al.~\cite{marchal2018parallel} assume a memory model where each processor has an individual memory available.
    Workflow tasks have no memory requirements, but they have input and output files that need to be stored in the memory.
    A polynomial-time algorithm for computing the peak memory needed for a parallel execution of such a DAG is provided,
    as well as an ILP solution to the scheduling problem.
    The memory model requires deleting all input data upon starting of the task and adding all output files there.

    In an assumed dual-memory systems~\cite{herrmann2014memory},  a processor can have access 
    to a memory of two different
    kinds (red or blue), and each task can be executed on only one sort of memory.
    The communications happen only between these two kinds of processors (communications within 
    each group are ignored).
    The authors then formulate an ILP solution for this problem formulation.

    %The algorithm presented by Yao et al.
    Yao et al.~\cite{yao2022memory} consider that each processor has an own internal memory and all
    processors share a common external one. The internal (local) memory is used to store the task files.
    The external memory is used to store evicted files to make room for the execution of a task on a processor.
    All processors, including the original one, can access these files. 
    Each edge %in~\cite{yao2022memory} 
    has two weights -- the size of the files transferred along it,
    and the time of communication along this edge.
    The tasks themselves have no memory requirements, but need to hold all their incoming and outgoing files.

    In~\cite{ding2024ils}, there are connected processors with individual limited memories.
    The collective set of memories forms the global memory, to which each processor has access, however the access time
    to global memory is different.
    Each memory access in the graph is modeled as a memory access token on the task, while the edges have no weights.
    The solved problem is how to allocate initial input data in processor memories so that the overall
    execution is minimized and the memories are not exceeded.
    The authors propose an integer linear programming model.
    %that minimizes the length of the critical path, including a greedy initial solution.

    In~\cite{rodriguez2019exploration}, the authors assume memory requirement on tasks represented as tiles.
    Each processor has individual memories to process the task, but only the shared memories store the tiles containing
    memory tiles occupied by memory tiles.

   Finally, there are some cloud-oriented models that include costs associated with memory usage~\cite{liang2020memory}.

   Overall, there are a variety of memory models, but, to the best of our knowledge, the only study on a multiprocessor
   platform that is fully heterogeneous, with individual memories, is the one from~\cite{DBLP:conf/icpp/KulaginaMB24},
   but where a partition of the workflow is proposed, hence preventing processor reuse. 
   Hence, in ~\cite{DBLP:conf/icpp/KulaginaMB24}, there is no need of communication buffers to store data 
   that should be communicated
   between two processors when tasks are ready to execute.

    \subsection{Dynamic/adaptive algorithms}

We now review related work in a dynamic setting. With no variation in task parameters, 
    DVR HEFT~\cite{SANDOKJI2019482} rather considers that new tasks arrive in the system. 
    They use an almost unchanged HEFT algorithm in the static step, executing three slightly
    varying variants of task weighting and choosing the variant that gives the best overall makespan.
    In the dynamic phase, they receive new tasks and schedule them on either idle processors or 
    those processors that give them
    the earliest finish time.
    %Task failures are not covered.

    Rahman~\etal~\cite{rahman2013}'s dynamic critical path (DCP) algorithm for grids maps tasks to machines
    by calculating the critical path in the graph dynamically at every step.
    %For all tasks they compute the earliest start time and absolute latest start time that are upper and lower bounds
    %on the start time of a task (differing by the slack this task has).
    %All tasks on this critical path have the same earliest and latest start times, because they cannot be delayed.
    They schedule the first task on the critical path to the best suitable processor and recompute the critical path.
    %The algorithm takes the first unscheduled task on the critical path each time and maps it on a processor identified for it.
    %If processors are heterogeneous, then the start times are computed with respect for the processor, and the minimum
    %execution time for the task is chosen.
    The heuristic also uses the same processor to schedule predecessor and successor tasks, as to avoid data transfer between processors.
    The approach is evaluated on random workflows of the size up to 300 tasks.

    Garg~\etal~\cite{GARG2015256} propose a dynamic scheduling algorithm for heterogeneous grids based on rescheduling.
    The procedure involves building a first (static) schedule, periodic resource monitoring and rescheduling the remaining
    tasks.
    The resource model contains resource groups (small tightly-connected sub-clusters), connected between each other.
    For each resource group, there is an own scheduler, and an overall global scheduler responsible for distributing
    tasks to groups.
    The static heuristic is HEFT with earliest start time as priority.
    Upon rescheduling, a new mapping is calculated from scratch, and this mapping is accepted if the resulting makespan
    is smaller than the previous one.
    The experiments were conducted on a single workflow with 10 tasks.
%
    % The authors define the execution time, estimated start time, data ready time,a dn estimated finish time per task.
    %The runtimes of tasks depend on processor speeds, are calculated in advance and stored in tables.

    %The algorithm first computes bottom levels for all tasks (execution time is average of all possible execution times).
    %THe bottom level represents the priority of the task, and tasks are sorted according to these priorities.
    %They then go through tasks and map than to such processors that minimize the earliest start times of this task's
    %successors.
    %To do this, the authors calculate the earliest finishing time of the task across all ressources, along with the
    %average communication and computation costs fir the dependent tasks.
    %
    %The rescheduling is triggered when either a load on a resource increases over a threshold, or if a new resource
    %is added.

    Most dynamic or adaptive algorithms are formulated for clouds, where the execution environment is not fixed,
    but constrained by cost.

    Wang et al.~\cite{wang2019dynamic} propose a dynamic particle swarm optimization algorithm to schedule workflows in a cloud.
    Particles are possible solution in the solution space.
    However, the dynamic is only in the choice of generation sizes, not in the changes in the execution environment.
    Similarly, Singh et al.~\cite{singh2018novel} addresses dynamic provisioning of resources with a constraint deadline.

    De Olivera~\etal~\cite{de2012provenance} propose a tri-criteria (makespan, reliability, cost) adaptive scheduling algorithm
    for clouds.
    They solve a set of linear equations that represent the cost of an execution based on the criteria.
    The authors test out 4 scenarios - one preferring each criteria, and a balanced one.
    The algorithm chooses the best virtual machine for each next task based on the cost given by the model.
    The authors used workflows with less than 10 tasks, but repeated them so that the execution had up to 200 tasks.
    %They do not report the runtime of the scheduling algorithm, only the speedup and cost saving it produces.
 %   The authors use provenance data to make scheduling decisions.

    Daniels et al.~\cite{daniels1995robust} formalize the concept of robust scheduling with variable processing times
    on a single machine.
    The changes in runtimes of tasks are not due to changing machine properties, but are rather task-related (that means
    that these runtime changes are unrelated to each other).
    The authors formulate a decision space of all permutations of $n$ jobs, and the optimal schedule in relation to a
    performance measure $\phi$.
    Then they proceed to formulate the Absolute Deviation Robust Scheduling Problem as a set of linear constraints.

While several related work consider building a new schedule once some variation has been observed,
we are not aware of work implementing a real runtime system that interacts with the scheduler,
and tested on workflows with thousands of tasks, as we propose in this paper. Furthermore, 
we are not aware of any previous work discussing dynamic algorithms combined with memory constraints. 
%\AB{Say why our approach in dynamic setting is different and novel}

%    \subsection{Other notable works}
%
%    \cite{palis1996task} present a clustering-based scheduling algorithm for a parallel execution and prove its quality.
%    They utilize task duplication when creating the clusters (grains).
%    Their scheduler then maps clusters to processors.
%    They assume unlimited processors with the speed of 1.
%    For each task, they compute the earliest starting time and find a cluster, where this tasks's start time is as close
%    to it as possible.
%    %The cluster growing algorithm adds one task to the cluster at a time, by adding tasks in nondecreasing order of
%    %release times.
%
%    GRASP (generally randomized adaptive search procedure)~\cite{feo1989probabilistic} conducts a number of iterations
%    to search for an optimal solution for mapping tasks on machines.
%    A solution is generated at each step, and the best solution is kept at the end.
%    The search terminates when a certain termination criterion is reached.
%    It generates better results than other algorithms, because it explores the whole solution space.
%
%    Avanes~\etal\cite{avanes2008adaptive} present a heuristic for networks in disaster scenarios.
%    These networks are a set of DAG-shaped scenarios, out of which one needs to be executed.
%    The scenario contains AND- and OR-branches, where AND-branches indicate activities that need to be executed in parallel.
%    The heuristic first determines similar activities and groups them together.
%    Then they allocate these groups to disaster responders and tasks within this group according to a constraint system.
%    The dynamic part deals with changes and distinguishes between retriable and compensation activities.
%    The heuristic calculates a new execution path with these tasks.
%
%
%    \cite{lutke2024hetsim} is a scheduling simulator that models heterogeneous software with memory and accelerator
%    (processor) speed heterogeneity.
%    Each accelerator has its own memory that can be zero.
%    Each accelerator's characteristics depend on the task it runs and are not fixed.
%
%
%
%
%    \cite{meng2018traffic} investigate scheduling on multi-core chips.
%    Their model is far from ours.
%
%
%    An online scheduling algorithm~\cite{Witt2018POS} assumes a DAG-structured workflow and learns task characteristics.
%    They prioritize tasks that have failed before or are well-predictable.
%
%

\section{Model} %: \skug{Full: 4, polished: 3}}
\label{sec:model}

%\skug{
%CHANGES IN THIS CHAPTER
%}
    We first describe the model for our target applications, which are (large scientific) workflows for which we do not have perfect a priori knowledge,
    in Section~\ref{sec.mod.work}.  Next, we define the execution
    environment, a heterogeneous system (in terms of processor speed and memory size),
    in Section~\ref{sec.mod.plat}. Finally, we present the optimization problem in
    Section~\ref{sec.mod.pb}. The key notation is summarized in Table~\ref{tabnotation}.

    \subsection{Workflow}
    \label{sec.mod.work}
    A workflow is modeled as a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
    $E$ is a set of directed edges of the form $e=(u,v)$, with \mbox{$u,v\in V$}, expressing precedence constraints between tasks.
    Each task~$u \in V$  is performing $w_u$ operations, and it also
    requires some amount of memory to be executed, denoted as~$m_u$.
    Each edge $e=(u,v) \in E$ has a cost~$c_{u,v}$ that corresponds to the size of the output file written by task~$u$ and used as input by task~$v$.

    Note that $m_u$ is the total memory usage
    of a task during its execution, including input and output files currently being read and written,
    and hence it is greater than the total size of input files
    (received from the predecessor tasks) and than 
    the total size of  output files (sent to successor tasks): 
    $$   m_u \geq \max \left\{ \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v} \right\} . $$

%     the total memory requirement  for the execution of task~$u$ consists of the maximum
%    between the input files
%    (total size of the files to be received from the parents),
%    the output files (total size of the files to be sent to the children),
%    and the total memory size~$m_u$ (usually achieving the maximum):
%    \[
%        r_u = \max\left\{m_u , \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v}\right\}.
%    \]

    The predecessors of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of predecessors is
    $ \parents{u} = \{v \in V: (v,u) \in E\}$. A task without predecessors is called a {\it source task}.
    The successors of a task~$u$ are the tasks following~$u$ directly according to the precedence constraints, i.e.,
    $ \children{u} = \{v \in V: (u,v) \in E\}$. A task without successors is called a {\it target task}.
    Each task may have multiple predecessors and successors.

Furthermore, we place ourselves in a context where we do not have perfect knowledge
of the length of the tasks, i.e., $w_u$, % ($w_u$ and $m_u$) 
before the tasks complete their execution,
but only estimates~\cite{rahman2013,GARG2015256}.  
%. \AB{Add motivation: related work with variable task durations for instance...}
Hence, scheduling decisions are made on these estimated parameters, and
may be reconsidered at runtime when a task completes its execution. 
%when a task starts its execution and we know its exact parameters.
%\AB{Variability only on $w_u$ for now}

    \subsection{Execution environment}
    \label{sec.mod.plat}

%\skug{changes here}
The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
consists of $k$ processors $p_1, \dots, p_k$.
Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$ and a speed~$s_j$.
All processors have access to a shared disk of unlimited size, where they can write
and read data: Processor~$p_j$ has a bandwidth $\bw_j$ (resp.~$\br_j$) to write  (resp. read),
and hence it takes a time $\frac{c_{u,v}}{bw_j}$ to write the data produced by task~$u$
on~$p_j$ for task~$v$, 
if this data is evicted from memory.

%, but of much slower speed.
All communications happen over the disk -- after one processor has finished writing data there, another one can read
it to load the data on its individual memory. 
Hence, we can decide to evict some data from the individual processor memories if it has been 
written onto the disk, 
in order to free some space in the individual memories.
This data can later be read back into memory of the same or any other processor at any time.
%\skug {check this statement: If data has been written on disk, it is removed from the local memory (no two copies can be kept at the same time).} \AB{We can keep two copies: evict if we need to }

If the entire memory required by  task ~$u\in V$ fits in the available memory of processor~$p_j$,
    then the execution time of this task on this processor is expressed as $\frac{w_u}{s_j}$.
However, if the memory requirement of the task exceeds the available memory on the processor, 
then task can still be executed there, but it will be slowed down. Indeed, 
%    albeit slower.
the part of memory requirement that exceeds the available memory on processor~$j$, $availM_j$, 
will be offloaded to the disk, slowing down the execution:

    \[
        w_{real} = \frac{w_u}{s_j} +  q_s \times   (m_u - availM_j) , %   \frac{q_s \times w_u}{s_j} \frac{m_i - M_j}{m_i}
    \]
where $q_s$ is the slowdown coefficient, and $(m_u - availM_j)$ is the part 
of the task memory that needs to be offloaded.
\AB{Svetlana please check, I think we said in the end that we use absolute value 
of the amount of memory not fitting locally...}
%    \skug{check this equation}
%We assume that all processors are connected with the same bandwidth~$\beta$.
%    \hmey{Maybe mention that variable bandwidths are part of future work...?}

Multiple processors can communicate with the disk in parallel, but the communication of each processor  with the disk is sequential.
The processor can read and write at the same time, but only one file can be read and only one written at any point in time.
We keep track of the current ready times of each processor $j$, ready time for computation $\rt_j^c$ (when one task has finished
computing and another one can start),
ready time for writing $\rt_j^w$ (when a file has been written to disk and another file can be written), and ready time for
 reading~$\rt_j^r$.
Initially, all the ready times are set to~$0$.
We also keep track of the currently available memory, $availM_j$ on the processor memory.

Furthermore, $\PD_j$ is a priority queue with the {\em pending data}
that are in the memory of processor~$p_j$,  but that may be evicted to the disk
if more memory is needed on~$p_j$.
They are ordered by non-decreasing size and correspond to some files~$c_{u,v}$.
When scheduling a task, the decision has to be made to execute this task on a processor with enough memory
without slowdown, but with possibly more time to read the input files, or to execute the same task
on a processor with less memory (and hence a slowdown), but potentially with a higher processor speed 
or with less times spent reading files.

We use the \algo{memDag} algorithm developed by Kayaaslan \etal~\cite{KAYAASLAN20181} to compute
the memory requirement; it transforms the workflow into a series-parallel graph
and then finds the traversal that leads to the minimum memory consumption.
\AB{When do we need this? Probably more in Algorithms, right? }

%\skug{Todo: describe deviations}

    \begin{table}
        \begin{center}
            \begin{tabular}{rl}
                \hline
                \textbf{Symbol}                       & \textbf{Meaning}                                         \\
                \hline
                $G = (V, E)$                          & Workflow graph, set of vertices (tasks) and edges        \\
                $\parents{u}$, $\children{u}$         & Predecessors of a task $u$, successors of a task $u$            \\
                $m_u$                                 & Memory weight of task $u$                                \\
                $w_u$                                 & Workload of task $u$  (normalized execution time)          \\
                $c_{u,v}$                             & Communication volume along the edge $(u,v)\in E$         \\
                $F$, $\mathcal{F}$                    & A partitioning function and the partition it creates     \\
                $V_i$                                 & Block number $i$                                         \\ %\wrt~some $F$   \\
                $\cluster$, $k$                    & Computing system and its number of processors           \\
                $p_j$, proc($V_i$)                          & Processor number $j$, processor of block $V_i$                 \\
                $M_j$, $MC_j$, $s_j$                               & Memory size, comm. buffer size, and speed of proc.\ $p_j$                          \\
                $\beta$                     & Bandwidth in the compute system                                \\
                $\bottomlevel{u}$                      & Bottom weight of task $u$ \\
                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$               \\
                $\Gamma = (\mathcal{V}, \mathcal{E})$                      & Quotient graph, its vertices and its edges        \\
                $r_u$, $r_{V_i}$                            & Memory requirement of task $u$ and of block $V_i$                 \\
                \hline
            \end{tabular}
        \end{center}
        \caption{Notation} \label{tabnotation}
    \end{table}

\subsection{Optimization problem}
\label{sec.mod.pb}

The goal is to find a schedule of the DAG~$G$ for the $k$ processors,
so that the makespan (total execution time) is minimized. 
\AB{no more memory constraint, now the lack of memory just slows down the execution, correct?
I'll probably add a few sentences here...}
% while respecting memory constraints. \skug{do we still formulate it like that?}
    %If a processor runs out of memory to execute
%a task mapped on it, the schedule is said to be {\em invalid}.

Since tasks are subject to variability, we aim at minimizing the actual makespan
achieved at the end of the execution, while decisions may be taken building
on the estimated task parameters.
\AB{Probably need to clarify offline vs online problem at some point...}

Note that the problem is already NP-hard even in the homogeneous case and 
without memory constraints, because of the DAG structure of the application. 
Hence, we focus on the design of efficient scheduling heuristics. 
%\hmey{Mention NP-hardness due to being more general than NP-hard problem?}

%    \paragraph{Workflow-related changes}
%
%    \begin{itemize}
%        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
%        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.
%
%    \end{itemize}
%
%    The following changes are not a part of this article's scope:
%
%    \begin{itemize}
%        \item The workflow structure changes: edges or tasks come in or leave.
%    \end{itemize}
%
%    \paragraph{Execution environment-related changes }
%
%
%    \begin{itemize}
%        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
%        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.
%
%    \end{itemize}
%
%    The following changes are not a part of this article's scope:
%
%    \begin{itemize}
%        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
%    \end{itemize}
%
%    \subsection{Time of changes }
%
%    We consider discrete time in seconds.
%    The time point(s) at which the changes happen is unambiguously defined.
%
%    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
%    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
%    The start time of the source task in the workflow is zero.
%    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$
%
%    \subsection{Changes and knowledge horizon - important questions TBA}
%
%    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.
%
%    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
%    However, how do we model the following:
%    \begin{itemize}
%        \item Do we know the new weights of currently running tasks and tasks that have not yet started? This means, do we foresee into the future or do we assume that all weights on unfinished tasks remain the same?
%        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
%        \item How many times of change do we model - one per workflow run, or multiple?
%        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
%    \end{itemize}

\section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
\label{sec:heuristics}
\subsection{\skug{New scheduling approach - general description}}

    In the first step, we order the tasks by non-increasing ranks (e.g. their bottom levels).
    In the next step, we repeatedly schedule ready tasks.
    For each task, we first choose the best processor to put it on - the one that minimizes the expected finishing time.
    For each processor, we compute this earliest finishing time by first calculating the earliest starting time on this processor.
    It is computed as the maximum of the ready time of computation on this processor, and the time to read input files.
    For the currently reviewed processor, we preliminarily schedule necessary file writes on other processors holding input files,
    and file reads on the current processor.
    Doing so, we respect the ready times on writes on other processors and reads on the current processor.
    We possibly plan ``into the past'' with ready and writes.

    The earliest finish time is calculated by adding the execution time (possibly with delay due to memory overflow) to the start time.
    Further scheduling decisions happen when another task finishes.
    \skug{check this: when a task finishes earlier or later than planned, we review our earlier scheduling decisions and may
    move the task execution to another processor if this proves more beneficial.}

    When a task finishes, we look at all of its sucessors, rather than only the ready ones.
    For tasks that lie further in the future, we try to plan rather early.
    Therefore, we employ lazy writes.
    We try to write the files for these tasks to the disk early.
    So, we preliminarily schedule these file writes. However, they are not high priority and can be cancelled.
    We keep two ready times on writes, one soft, with lazy writes, and another hard onw, with scheduled writes.
    If a task requires a write during the lazy write, we cancel it and move it to later point in time.

    \skug{For each edge, we also keep information on where it is currently stored - in memory of a processor (with processor id),
        on the disk or nowhere - if this file has not yet been generated or has already been consumed by the successor task.}

\subsection{Data structures }
    To optimally represent the timeline of the execution of a workflow, we represent the execution as a series of events
    with attached unique timestamps.
    The events are kept in the priority queue ordered by timestamp.
    Events are created when other events are triggered and are being inserted into the queue.
    When their timestamp is the smallest, they are then triggered, possibly creating new events.
    The execution starts with placing the event of starting the starting node into the queue with the timestamp 0.
    The execution ends when there are no events in the queue.

    An event can happen on finish or start of a task execution or an io action (file read or write).
    Each event has several parameters:
    \begin{itemize}
        \item Event type: OnTaskStart, OnTaskFinish, OnIoStart, OnIoFinish
        \item The workflow entity associated with the event: the task to be executed or an edge whose file needs to be read or written
        \item The processor id on which the action should take place
        \item Expected and actual time of firing.
        \item A set of predecessor events. The event is called ready if all its predecessors have been finished.
        \item A set of successor events to keep track of.
        \item For a write event, a boolean if this file should be removed from memory after being written.
    \end{itemize}

    Per processor,  in addition to keeping the fixed values (memory, speed), we keep track also of the current state in its memory.
    Because during and after execution of each even the state of the processor memory changes, we need to keep snapshots of what files
    were in memory and how much of it was available.
    Because keeping such snapshots for each timestap is too complicated, we only keep two timestamps.
    First, we keep the information of available memory and tasks pending in memory during the execution of the last
    task scheduled to this processor.
    Then, we keep the same information on what it will be immediately after this task finishes.

    \subsubsection{On Task Finish}
    When a task finishes, it first cleans after itself:
    \begin{itemize}
        \item It frees its memory: it increases the available memory on the processor by the amount it occupied
        \item If there had been a memory overflow onto disk, then this memory is being freed implicitly
        \item it sets its status to finished in the workflow.
    \end{itemize}
    It the goes over its successor events (Tasks and IO actions) and removes itself from their predecessor lists.
    It also updates their actual trigger times with its actual finish time.
    Successor tasks can be tasks scheduled on the same processor after this one.
    Successor io events can be reads that could not start until the task finishes and frees its memory.
    \skug{can writes be successor events of a task?}

    Then task goes over its successor tasks in the workflow and schedules ready ones.
    Ready tasks are those whose all predecessors have the finished status.

    For each ready task, the scheduler tries to tentatively put it on each processor in the cluster and chooses the processor
    that promises the earliest finish time.
    The finish time on each processor is calculated as follows:

    \paragraph{Earliest start and finish time}
    We find out when the last task on this processor finishes.
    If there is enough memory to start our task after that, then the end of previous task is the earliest start time for our task.
    If files need to be evicted first, we need to find a balance between evicting tasks (and waiting for their writes to finish),
    then running our task, and starting the task immediately and letting it run longer due to memory overflow.
    We compare the following cases:
    \begin{itemize}
        \item Evict everything, then start task
        \item Greedily evict the largest file, starting the task
        \item Not evicting anything, starting the task
    \end{itemize}
    Whatever combination gives the best estimated finishing time for the task, we note the estimated starting time of the task in this case.
     The finish time is the earliest start time  plus computation time of the task.

    \paragraph{Incoming edges} For each incoming edge, its location is determined.
    If the edge is in the memory of the target processor, nothing needs to be done.
    The estimated start and finish times for the task are unaffected.

    If the edge is on the disk, it first needs to be read into the memory of the target processor.
    We delay the reads to avoid holding unnecessary files in the memory.
    When we preliminarily schedule a read, we first assess its length and other actions that happen on this processor in the meantime.
    So, the first estimated start of the read is the estimated start time of the task minus the estimated length of the read.
    We then see if this start of the read is happening during the runtime of the previous task or already after it (during the writes).
    If the starting time of the read is during the execution of the previous task and there is enough memory to read this file, then we move the read
    forward to the beginning of this previous task.
    If there is not enough memory, then the entire read is rescheduled for after the task finishes.
    The read will then be bottleneck and affect the eraliest start time of the desired task.

     If the edge is in another processor's memory, it needs to be first written to disk, then read by the target processor.
    So we look at the latest write on that processor, preliminarily schedule the write to disk and then proceed like in the
    previous case, reading the file from disk.

   \paragraph{When processor is chosen} When a processor is chosen, the scheduler inserts new events into the queue:
    \begin{itemize}
        \item For the task, OnTaskStart and OnTaskFinish events with the estimated trigger times, the OnTaskFinish event being dependent on the OnTaskStart event.
        \item For each io operation necessary for the execution: OnIoStart event of the corresponding type and on corresponding processor
        \item It connects all the event with predecessor-successor relations.
        \item If files needed to be evicted
    \end{itemize}
    The scheduler also updates the ready times of the processors and updates the new currently last events there.

    \subsubsection{On Task Start}
    When a task start event is triggered, it first determines if it can start.
    If some predecessor tasks have not yet finished (the predecessor set is not empty), then the actual time of this event
    is set to the maximum of the estimated end times of predecessor events and this event is reinserted in the queue
    with this new time.

    If the task can start, the scheduler determines the deviation of the runtime of the task.
    It applies the deviation function on its estimated runtime (estimated finish minus estimated start time).
    It then adds this deviated runtime to its actual start time and receives an updated actual finish time.
    With this time, it updates the OnTaskFinish event of this task.

    \subsubsection{OnIoFinish}
    When the io operation finishes, it removes itself from the predecessor sets of all its successor events.
    It also sets their actual time to its own actual finish time.

    \subsubsection{OnIOStart}
    When the IO operation starts, it updates the corresponding IoFinish event

    After being triggered and completed, each event is deleted from the queue.

--------------------------------------

%    The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
%    but favor processor reuse as is done in HEFT.
%    Furthermore, this would allow us to handle variability on the fly, by updating
%    the bottom levels if some parameters vary, and computing the schedule
%    only for the near future...

%In order to be able to easily adapt to variability of task parameters, 
We design 
variants of HEFT that account for memory usage and aim at minimizing the makespan.
First, we present in Section~\ref{sec.heft} the baseline HEFT heuristic that does not account for the memory
(and hence, may return invalid schedules that will not be able to run on the platform
by running out of memory).  Then, Section~\ref{sec.heftm} focuses on the presentation of the novel
heuristics, including eviction strategies to move some data in communication buffers
in case there is not enough memory available on some processors.

    \subsection{Baseline: original HEFT without memories}
\label{sec.heft}

    Original HEFT does not consider memory sizes.
    The solutions it provides can be invalid if it schedules tasks to processors without sufficient memories.
    However, these solutions can be viewed as a ``lower bound'' for an actual solution that considers memory sizes.

    HEFT works in two stages.
    In the first stage, it computes the ranks of tasks by computing their non-increasing bottom levels.
    The bottom level of a task is defined as
    $$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
    (the max is 0 if there is no outgoing edge).
    The tasks are sorted by non-decreasing ranks.

    In the second stage, the algorithm iterates over the ranks and tries to assign the task to the processor where it
    has the earliest finish time.
    We tentatively assign each task~$v$ to each processor~$j$.
    The task's starting time $st_v$ is dictated by the maximum between the ready time of the processor~$rt_j$,
    and all communications that
    must be orchestrated from predecessor tasks $u\notin T(p_j)$.
    The starting time is then:

 {\footnotesize{   \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]}}

    Finally, its finish time on $p_j$ is
    $FT(v,p_j) = st_v + \frac{w_v}{s_j}$.

    Once we have computed all finish times for task~$v$,
    we keep the minimum $FT(v,p_j)$ and assign task~$v$
    to processor~$p_j$.

    \textit{Assignment to processor. }
    When assigning the task, we set the ready time $rt_j$ of  processor~$j$ to be the finish time of the task.
    For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
    communication buffers $rt_{j', j}$ for every predecessor $u$'s processor $j'$: we increase them by the
    communication time $c( u,v) / \beta$.

    \subsection{Memory-aware heuristics}
    \label{sec.heftm}
    Like the original HEFT, the memory-aware versions of HEFT consist of two stages:
    first, they compute the task ranks,
    and second, they assign tasks to processors in the order defined in the first stage.
    We consider three variants of HEFT accounting for memory usage (HEFTM), which only
    differ in the order they consider tasks to be scheduled in the first stage.

\smallskip
    \noindent{\bf Compute task ranks. }

We design three variants of memory-aware HEFT:
\begin{itemize}
\item    HEFTM-BL orders tasks by non-increasing bottom levels, where the bottom
    level is defined as
    $$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
    (the max is 0 if there is no outgoing edge).

\item    HEFTM-BLC %: from the study of the fork (see below), it seems important
   %  to also account for the size of the data as input of a task,
    is  giving more priority at tasks with potential large incoming communications,
    hence aiming at clearing the memory used by files as soon as possible,
    to have more free memory for remaining tasks to be executed on the processor.
    Therefore, for each task, we compute a modified bottom level accounting for communications:
    $$blc(u) = w_u + \max_{(u,w)\in E} \{c_{u,w} + blc(w)\} + \max_{(v,u)\in E} c_{v,u}   . $$

%    \skug{avoid having mixed ranks, when the memory size of the lower task is not taken into account}

\item   Finally, HEFTM-MM orders tasks  in the order returned by %as dictated by MinMem.
the  \algo{memDag} algorithm~\cite{KAYAASLAN20181}, which corresponds to a traversal
of the graph that minimizes the peak memory usage.

\end{itemize}

\smallskip
  \noindent  {\bf Task assignment. }

    Then, the idea is to pick the next free task in the given order,
    and greedily assign it to a processor, by trying all possible options
    and keeping the most promising one. We first detail how a task
    is tentatively assigned on a processor, by carefully accounting for the memory usage.
    Next, we explain the steps to be taken to effectively assign a task on a given processor.

    \medskip
    \noindent{\em Tentative assignment of task~$v$ on $p_j$.}\\
    {\bf Step 1.} First, we need to check that for all predecessors~$u$ of~$v$ that are mapped
    on~$p_j$, the data $c_{u,v}$ is still in the memory of~$p_j$,
    i.e., $c_{u,v}\in PD_j$. Otherwise, the finish time is set to~$+\infty$ (invalid choice).

    \smallskip
    \noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
    $$Res = availM_j - m_v - \!\!\!\! \sum_{u \in \Pi(v), u\notin T(p_j)}  \!\!\!\!\!\!\!\!\{c_{u,v}\}
    - \sum_{w\in Succ(v)} \!\!\!\!\!\! \{c_{v,w}\}.$$

    $T(p_j)$ is the set of tasks already scheduled on $p_j$; by step 1, their files are
    already in the memory of~$p_j$. However, the files from the
    other predecessor tasks must be loaded in memory before executing task~$v$,
    as well as $m_v$ and the data generated for all successor tasks.
    $Res$ is then checking whether there was enough memory; if it is negative,
    it means that we have exceeded the memory of~$p_j$ with this tentative
    assignment.

    In this case ($Res <0$), we try evicting
    some data from memory so that we have enough memory to execute task~$v$.
    We need to evict at least $Res$ data.
    We propose a greedy approach, evicting the largest files of $\PD_j$ until $Res$ data has been evicted. 
    A variant where the smallest files are evicted first has been tested, and it led to comparable results. 
%    
%    in order to avoid costly communications.
%    \AB{FYI We initially discussed evicting the largest files, but this leads to
%    large communications and does not seem efficient after all... Maybe we can think of another
%    approach that would take into account both data size and bottom level...}
    While tentatively evicting files, we remove them from the list of pending memories and move them into a list
    of memories pending in the communication buffer.
    We keep track of the available buffer size, too -- each time a file gets moved into the pending in buffer, 
    the available buffer size is reduced by its weight.

    If we still do not have enough memory after having tentatively evicted all files from $\PD_j$,
    or if while doing so we exceeded the size of the available buffer,
    we set the finish time to~$+\infty$ (invalid choice).

    \smallskip
    \noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$.
    Its starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
    must be orchestrated from predecessor tasks $u\notin T(p_j)$.
    The starting time is therefore:\\[-.7cm]

   {\footnotesize{ \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v), u\notin T(p_j)}\{ FT(u) , rt_{proc(u), p_j}\} + c_{u,v} / \beta \} .} \]}}

  \noindent  Finally, its finish time on $p_j$ is 
    $FT(v,p_j) = ST(v, p_j) + \frac{w_v}{s_j}$.

    \medskip
    \noindent{\em Assignment of task~$v$.}\\
    Once we have computed all finish times for task~$v$,
    we keep the minimum $FT(v,p_j)$ and assign task~$v$
    to processor~$p_j$.
    In detail, we:
    \begin{itemize}
        \item  Evict the file memories that correspond to edge weights that need to be evicted to free the memory.
        We remove these files from pending memories
        $PD_j$, add them to pending data in the communication buffer, and reduce the available buffer size accordingly.
        \item    Calculate the new $availM_j$ on the processor.
        We subtract the weights of all incoming files from predecessors assigned to the same processor,
        and add the weights of outgoing files generated by the currently assigned task.
        \item  For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
        communication buffers $rt_{j', j}$ for the processor~$j'$that the predecessor $u$ has been assigned to: we increase them by the
        communication time $c( u,v) / \beta$.
        We also remove the incoming files from either the pending memories or pending data in buffers of these other
        processors, and increase the available memories or available buffer sizes on these processors.
        \item We compute the correct amount of available memory for~$p_j$ (for when the task is done).
        Then, for each predecessor that is mapped to the same processor, 
        we remove the pending memory corresponding to the weight of
        the incoming edge, also freeing the same amount of available memory (increasing $availM_j$).
        For each successor, we rather add the edge weights to pending memories and reduce $availM_j$ 
        by the corresponding amount.
    \end{itemize}

\hmey{BEGIN FORK}
    \subsection{The fork}
    We look at the behavior of these heuristics on a fork graph,
    where there is a root task~$T_0$, producing $n$ files $f_1, \ldots, f_n$
    to be used by tasks $T_1, \ldots, T_n$ ($f_i = c_{0,i}$).

    Without memory, this problem is NP-complete; this is equivalent
    to 2-partition if the tasks have $w_i=a_i$, and all files are of size~$f_i=0$,
    and with two processors. Half of the tasks must be sent to the processor
    on which $T_0$ is not executed, and the optimal makespan is
    $w_0+\frac{1}{2}\sum_{1\leq i \leq n} w_i$.

    However, with an infinite number of identical processors, it can be
    solved in polynomial time: sort tasks by non-decreasing $f_i+w_i$;
    the $k$ tasks with smallest $f_i+w_i$ are then sent to another processor,
    while the remaining $n-k$ tasks are executed locally (try all values of $k$).

    With heterogeneous processors, it is probably NP-complete again
    because we could ensure that there are only two processors fast enough
    and get back to the 2-partition...

    We also had an example where evicting large files first in step 2
    can lead to arbitrarily bad makespan. Consider a fork with $n=2$,
    $f_1=1$, $w_1=2$, $f_2=100$, $w_2=1$, and memory constraint
    imposes that we free one unit of memory before executing one
    of the tasks\ldots Actually the new version with BLC would start
    considering $T_2$ and be fine in this case\ldots

\begin{figure}[tb]
  \centering
  \includegraphics[width=0.5\columnwidth]{images/bad-blc-instance.png} \\[1ex]
  \includegraphics[width=0.999\columnwidth]{images/blc-bad-cost.png} \\[1ex]
  \includegraphics[width=0.58\columnwidth]{images/opt-cost.png}
  \caption{Top: Example instance $I$, middle: BLC applied to $I$ leads to makespan $w_0 + nk/2 + k$, bottom: optimal solution has makespan $w_0 + n/2$.
  \hmey{TODO: adapt resolution of figure and fix $\epsilon$ problem}}
  \label{fig:fork-bad-blc-instance}
\end{figure}

Figure~\ref{fig:fork-bad-blc-instance} shows an instance where \heftblc\ performs very poorly.
Since the sorting criterion $w_i + c_{0i}$ yields an arbitrary ordering of the tasks $T_i$ (with $1 \leq i \leq n$), we may
assume that the algorithm processes the nodes in level $1$ of instance $I$ from left to right. 
As shown in the middle part of the figure, $T_1$ is processed...
\hmey{TODO: adapt with $\epsilon$ to enforce even tasks to be ``sent away''}
This leads to a makespan
of $w_0 + nk/2 + k$. The optimal solution, in turn, as shown in the bottom,
requires only $w_0 + n/2$ timesteps. Since $k$ can be chosen arbitrarily large if the value $10$ is scaled accordingly,
this instance shows that \heftblc\ may yield solutions with an arbitrarily large gap to optimality.

\AB{Can we prove that we have (maybe) a 2-approximation,
        at least for the fork? What worst-case can we think of? }
\hmey{END FORK}

\section{Dynamic scenario}
\label{sec:dyn}

    In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.
    This information may include memory usage, runtime, graph structures, or the status of the underlying infrastructure.
    In order to ensure that the information is up to date, a monitoring system observes the workflow execution and collects metrics for tasks and the underlying infrastructure.
    By incorporating dynamic monitoring values, e.g., the resources a task consumed, the runtime environment can incorporate the data into the prediction model to provide more accurate resource predictions.
    Also the underlying infrastructure can change during the workflow execution.
    Examples are processor failures, node recoveries, or acquisition of new nodes.
    However, also when the hardware of the infrastructure does not change, the set of nodes provided as a scheduling target might change due to release or occupation in shared cluster infrastructures.
    As infrastructure information and resource predictions are dynamically updated and provided to the scheduler during the workflow runtime, the previous schedule becomes invalid and a new one must be calculated.

    For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures
    of approximately 15\% is shown~\cite{malik2013execution}.
    Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one-third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
%For instance, Nadeen~et~al.\cite{} report an error of 10\%, 11\%, and 15\% while the task prediction errors shows a normal and exponential distribution.
%Bader~et~al.~ report a prediction error between 13\% and 17\% for their method, showing an exponential task error distribution.
% @Svetlana, willst du sowas für deine Experimente? Also die Daten, welche du dann konfigurieren kannst?
    Such a dynamic execution environment requires a dynamic scheduling method where the schedule can be recomputed during the workflow execution.

    \subsection*{Retracing the effects of change on an existing schedule}
    After the monitoring system has reported changes, we need to assess their impact on the existing schedule.
    These changes can invalidate the schedule (\eg if there is not enough memory for some tasks to execute anymore),
    they can lead to a later finishing time (\eg if some tasks are longer and they delay other tasks), or they can have no effect (\eg if new processors
    joined the cluster, but the old schedule did not account for them).
    To assess the impact, we need to retrace the schedule.

    First, we find out if at least one processor that had assigned tasks has exited - this instantly invalidates the
    entire schedule.

    We then iterate over all tasks of the workflow in a topological order - any of the orderings given by rankings BL, BLC or MM
    is a topological ordering.
    We then repeat steps similar to those we did during tentative assignment in the heuristics, 
    except that we do not choose a processor
    anymore, but rather we check whether the current processor assigned to the task still fits.

    For each task $v$, we first assess its current memory constraint $Res$ using Step 2 from the heuristic.
    The factors that affect $Res$ are possible changes in $m_v$, in $c_{u,v}$ from predecessors $u$ 
    or $c_{v,w}$ from successors $w$,
    available memory $availM_j$ on the processor (due to either changed $M_j$ or changed memory requirements 
    from other tasks).
    If originally, $Res$ was positive (no files were evicted from memory into the communication buffer), 
    then it has to stay this way --
    otherwise, evicted files can invalidate next tasks.
    If original $Res$ was negative, then we need to make sure that evicted files still fit into the communication buffer.
    If either $Res$ is newly negative, or the communication buffer is not large enough, this invalidates the schedule.
    We update the $availM_j$ and $availMC_j$ values according to the new memory constraints.

    Then, we can re-calculate the finish time of the task on its processor like in Step 3.
    The factors that affect it are changes in own execution time $w_v$ of the tasks, changed ready time of the processor
    (after delayed previous tasks), and changed communication buffer availability.

    Finally, after having updated the processor's values, we move on to the next task.

%    \subsection{Approximation}
%    \hmey{Rough notes:}
%    Let's use a fork to see how the algorithm behaves and if it provides some approximation. Our current intuition is that, if the memory constraint is ignored, HEFTM-BLc provides a $2$-approximation (to be proved).

\section{Experimental evaluation} %: \skug{Full: 3, Polished: 3}}
\label{sec:expe}

We first describe the experimental setup in Section~\ref{sec:setup}. 
Then, we report results on static experiments to assess the performance
of the memory-aware heuristics in Section~\ref{sec.expe.static}, before
discussing the heuristics behavior in a dynamic setting in Section~\ref{sec.expe.dyn}. 
Finally, we report running times of the heuristics in Section~\ref{sec.expe.t}.

\subsection{Experimental setup} %: \skug{Full: 5, Polished: 4}}
    \label{sec:setup}

    All algorithms are implemented in C++ and compiled with g++ (v.8.5.0).
    The experiments are managed by simexpal~\cite{DBLP:journals/algorithms/AngrimanGLMNPT19} and executed on workstations with 192 GB RAM and 2x 12-Core Intel Xeon 6126 @3.2 GHz
    and CentOS 8 as OS.
    Code, input data, and experiment scripts are available to allow reproducibility of results, at~\url{https://zenodo.org/records/13919214}~and~\url{https://zenodo.org/records/13919302}.

    We first describe the set of workflows used for the evaluation, and then the clusters on which the
    workflows are scheduled.

    \subsubsection{Workflow instances}
  %  \skug{Todo: describe the workflows the same asin the abstract}
    We run  experiments on a dataset that consists of real-world workflows: 
    we use workflows coming from~\cite{lotaru} (atacseq, bacass, chipseq,
    eager and methylseq), and also we increase their size using the WFGen generator~\cite{COLEMAN202216}.

    \paragraph{Workflow graphs}
    For the five real-world workflows, their nextflow definition (see~\cite{ewels2020nf}) was downloaded from the
    respective repository and transformed into .dot format using the nextflow option ``-with-dag''.
    The resulting DAG contains many pseudo-tasks that are only internal representations in nextflow
    (and not actual tasks); that is why we removed them.

    For the size-increased workflows, the graph is produced by the WFGen generator, based on a {\em model workflow} and
    the desired number of tasks.
    We used the real-world workflows as models, except for bacass that lead to errors in the generator.
%
    As number of tasks, we use: 200, \numprint{1000}, \numprint{2000}, \numprint{4000}, \numprint{8000}, \numprint{10000},
    \numprint{15000}, \numprint{18000}, \numprint{20000}, \numprint{25000}, \numprint{30000}.
    We divide the workflows into four groups by size: tiny ones with up to \numprint{200} tasks, small ones with \numprint{1000} to \numprint{8000} tasks,
    middle ones with \numprint{10000} to \numprint{18000} tasks, and big ones with \numprint{20000} to \numprint{30000} tasks.

    \paragraph{Task and edge weights}
    For the real-world workflows, we use historical data files provided by Bader~\etal~\cite{lotaru}.
    The columns in these files are measured Linux PS stats, acquired during an execution of a nextflow workflow.
    Each row corresponds to an execution of one task on one cluster node.
    Since the operating system cannot distinguish between (a) the RAM the task uses for itself and (b) the RAM it uses
    to store files that were sent or received from other tasks, the values in the historical data are total memory requirements (input/output files plus memory consumption of the computation).
    In a similar manner, the historical data provided by~\cite{lotaru} do not store the actual weights of edges between tasks, but only the overall
    size of files that the task sends to all its children.

    For each task, historical data can contain multiple values, obtained from the runs with different input sizes.
    The same workflow can require different memory capacity and take different time to execute
    depending on the size of its input.
    We simulate these various runs by obtaining values corresponding to each input size.
    For each of the four families, there are five input sizes, so we run each workflow in five variants corresponding to these inputs

    Not all tasks have historical runtime data stored in the tables.
    In fact, for two workflows, Bader~\etal do not provide data for more than 50\% of the tasks.
    For two more, around 40\% of tasks have no historical runtime data stored.
    Hence, in the absence of historical data about a task, we give it fixed weights.
    We give it an execution time of $1$, a memory requirement of $50 MB$, and files written and received of $1KB$.
    These values align with the findings of~\cite{lotaru} about small tasks.
%
 %   \skug{insert more abt how the runtime system hanfles data reads}

    \subsubsection{Target computing systems}
%    \skug{Because of the normalized nature of the task/edge weights, I will now be saying that our larger cluster has 192 GB and it
%    is represented as 192 000 (with 3 zeroes). For the memory-constrained cluster (192 00 with 2 zeroes) I will say that it has 10 times
%    less memory - 19.2 GB. Other processors will be dealt with similarly.}

    To fully benefit from the historical data, the  {\em default} experimental environment
    that we consider is a cluster based on the same six
    kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
    We set the number of each kind of node to $12$, thus having 72 processors in total.  % the whole cluster.
%
    Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as (name, speed, memory):
    ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines;
    ($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and
    large memory (see Table~\ref{tab:procs}).

    \begin{table}[htb]
        \begin{center}
            \begin{tabular}{c|c|cc}
                \toprule
                Processor  %& Amount
                &  CPU speed   & \multicolumn{2}{c}{Memory size (GB)} \\
                name & (GHz) & {\em default} & {\em mem-constrained} \\
                \midrule
                local                    & 4                    & 16     & 1.6 \\
                A1                      & 32                   & 32     & 3.2 \\
                A2                      & 6                    & 64    & 6.4 \\
                N1                      & 12                   & 16     & 1.6 \\
                N2                      & 8                    & 8      & 0.8\\
                C2                      & 32                   & 192   &  19.2\\
                \bottomrule
            \end{tabular}
        \end{center}
        \caption{Cluster configuration.}
        \label{tab:procs}
    \end{table}

    We also consider a more constrained setting, by varying the cluster configuration. The
     {\em memory-constrained cluster} is consisting of 72 nodes (12 of each kind) as the default cluster,
     but each node has 10 times less memory. Hence,  
     the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB, 
     $A2$ has $6.4$~GB instead, of $64$~GB etc.
    The processor speeds and their relations stay unchanged (see Table~\ref{tab:procs}).

Note that in both clusters, we set the size of the communication buffer to be equal
to ten times the the memory size.

    \subsubsection{Runtime system}

    To simulate the execution of a workflow, we implemented a runtime system.
    It reads the historical data and builds weights for tasks, as explained above.
%    These weights are then treated as the ``truth'' about the execution during the simulation.
    In the static case, these values are being sent to the scheduler, which builds a schedule
    according to these weights.
    However, in the dynamic setting, the runtime system applies a deviation function to the values.
    The deviation function computes a normally distributed deviated value, where the initial value
    is the mean and the deviation is~$10\%$.
    This scenario corresponds to the real-life scenarios identified in~\cite{lotaru} and
    other works dedicated to predicting runtimes of tasks~\cite{da2015online,da2013toward}.

    Hence, the scheduler receives deviated values and makes decision based on them.
    This leads to several types of possible issues:
    \begin{itemize}
        \item A processor is blocked by another task. If the scheduler underestimated the execution time of a task,
        it will block another one from starting.
        \item A predecessor has not yet finished.
        The scheduler may request a task to start on its processor, while
        some of the predecessors of the tasks have in fact not yet completed their execution, and the
        task is therefore not yet ready.
%        A similar situation, but a task could start on its processor, but one or more of its parents a
%        are not ready yet.
        \item Not enough memory. If the scheduler underestimated the amount of memory a task requires, this task might not be able to execute on a chosen processor.
        \item A task took less time than expected. We only consider this case if a task took more than 10\% less
        time than expected. In this
        case, we want to exploit the newly acquired free time by possibly starting other tasks earlier.
    \end{itemize}

\subsection{Results in a static setting} % \skug{Full: 3, Polished: 2}}
\label{sec.expe.static}

   % \subsubsection{Preliminary experiments}
We first study the heuristics behavior when the weights do not change upon runtime,
hence the scheduler has a perfect knowledge of task memory requirements and execution times.
We have compared two eviction strategies, starting with large files first or small files first,
and did not report any significant changes in terms of validity of schedule or makespan.
Hence, we only present results with the eviction of largest files first.
%    Our preliminary experiments have shown that evicting large or small files does not give a notable change in makespan.

    \subsubsection{Scheduling on the default cluster}

%    \paragraph{Normal cluster}

    On the default cluster,  the three memory-aware heuristics are able to schedule all workflows
    (see Figure~\ref{fig:success-rates-large}), while
    the baseline \heft has $24.2\%$ success rate ($75.7\%$ failure rate).
    Indeed, \heft is only able to schedule small workflows, but no workflow over the size of $4000$ tasks
    can be scheduled correctly; some tasks run out of memory. As soon as we are not in a setting
    with abundant processing resources for small workflows, it is hence necessary to adopt
    a memory-aware strategy in order to produce valid schedules.
%    Small workflows are less constrained by the memory requirement, so in the situation of abundant processing resources, even
%    a not memory-aware approach is able to produce valid schedules.
%    However, this advantage quickly vanishes as workflows grow larger.

    \begin{figure}[tb]
        \centering
        \includegraphics[width=0.9\columnwidth] {images/success-rates-large}
        \caption{Success rates on the default cluster. Higher is better.}
        \label{fig:success-rates-large}
        \vspace{-0.3cm}
    \end{figure}

   We also report in  Figure~\ref{fig:ms-relations-by-workflow}  the relative makespan found
   by the memory-aware heuristics, normalized to the
   makespan achieved by \heft, often through an invalid over-optimistic schedule that exceeds
   the bound on memory.
%    Figure~\ref{fig:ms-relations-by-workflow} shows the relations of the makespans found by the heuristics to the
%    makespans found by the baseline (including invalid ones), by workflow size.
    On average, the makespans found by \heftbl are $7.8\%$ worse than those found by the baseline,
    the makespans of \heftblc are $8\%$
    worse, and those found by \heftmm are $82.6\%$ worse. These are still very encouraging results,
    in particular for \heftbl and \heftblc, since the makespans of \heft correspond to invalid schedules.

 %   \skug{TODO: how much worse are makespans for SUCCESSFUL hefts? presulmably almost identical}

%    \begin{figure}[tb]
%        \centering
%        \includegraphics[width=1.1\columnwidth] {images/MsRelations2}
%        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline. Smaller is better.
%        \skug{SUGESTION: remove this picture, as it duplicates Fig 3}
%        }
%
%        \label{fig:ms-relations}
%        \vspace{-0.3cm}
%    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size}
        \caption{Relative makespans of heuristics normalized by \heft makespan, by workflow
        size, on default cluster. Smaller is better.}
        \label{fig:ms-relations-by-workflow}
        \vspace{-0.3cm}
    \end{figure}

%    \paragraph{Memory Usage}

    Finally, we study the percentage of memory occupied by the schedule, which is another good indicator
    of the memory usage of the heuristic and its ability to produce valid schedules.  %its important characteristic.
    Figures \ref{fig:mem-usages-normal} and~\ref{fig:mem-usages-onlyvalid} show the percentage of memory
    occupied on average by the schedule
    produced by the different heuristics, for different workflow sizes, first on all schedules (hence including
    \heft schedules that were not valid), and then only on valid schedules (hence, no results for \heft
    on large workflow sizes).
%
%   Figure shows memory usages if only valid heft schedules are presented.
%    There are no valid heft scheduler from middle-sized and large workflows, so no red bars are present.

    \heftmm continuously outperforms other heuristics in terms of memory usage,
    using from $46\%$ less memory on the smallest
    workflows to $4$ times less memory on the largest \numprint{30000}-task workflows.
    If we consider the invalid \heft schedules, too, we see that they would require more and more memory on average,
    which explains why these schedules rapidly become invalid.
    This is because some assignments require more that $100\%$ of memory (which makes them invalid).
    We can assess the degree of invalidity by comparing \heft memory usage with the memory usage of \heftbl.
    \heftbl differs from the baseline only in the sense that it respects available memory on the processors.
    For the largest workflows, \heft schedules require almost twice the memory of \heftbl.

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal}
        \caption{Memory usage on default cluster, including invalid \heft schedules. }
 %by algorithm by workflow size
         \label{fig:mem-usages-normal}
        \vspace{-0.3cm}
    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
        \caption{Memory usage on default cluster, including only valid \heft schedules. }
        \label{fig:mem-usages-onlyvalid}
        \vspace{-0.3cm}
    \end{figure}

\subsubsection{Scheduling on the memory-constrained cluster}

    On the memory-constrained cluster, \heft could produce valid assignments in only $14$ experiments
    out of $290$~($4.8\%$ success rate).
    The successful schedules were achieved exclusively on the tiny workflows (with only two 200-task size-increased
    workflow among them).
    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$ of them, while \heftmm
    could still schedule all the workflows, including even the largest ones, see Figure~\ref{fig:success-rates-tiny}.
    As also observed on the normal cluster, \heftmm seems to be less affected by the size of the workflow
    when scheduling it than the other heuristics.

    Similarly to the default cluster, we observe that relative makespan of \heftmm is usually greater
    than those of \heft (see Fig.~\ref{fig:ms-relations-by-workflow-constrained}), but \heft schedules are
    almost all invalid. It is therefore very interesting to resort to \heftmm for large workflows
    in a constrained cluster, since tasks are processed in an order that minimizes
    the memory usage of schedule.
%
%    The makespans it produces are larger, but it seems to order the vertices in a similar way on both the small and the large workflows,
%    rather than resorting to more memory-consuming ordering that other heuristics and the baseline produces.
%    This makes it especially valuable for very large workflows in situations of very constrained memory in the cluster.
%
%

    Memory usages on the constrained cluster are depicted in~Fig.~\ref{fig:memory-usage-constrained},
    and we observe that the memory footprint of \heftmm remains constant with workflow size.

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1\columnwidth] {images/success-rates-tiny}
        \caption{Success rates on the memory-constrained cluster. Higher is better. }
%        by algorithm by workflow size.
        \label{fig:success-rates-tiny}
        \vspace{-0.3cm}
    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1\columnwidth] {images/makespan_relations_by_wf_size-constrained}
        \caption{Relative makespans on the memory-constrained cluster.
%        produced by the heuristics to the makespan produced by the baseline, by workflow size.
        Smaller is better.}
        \label{fig:ms-relations-by-workflow-constrained}
        \vspace{-0.3cm}
    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/mem-usage-constrained-onlyvalid}
        \caption{Memory usage on the memory-constrained  cluster. } % by different heuristics by workflow size.}
        \label{fig:memory-usage-constrained}
        \vspace{-0.3cm}
    \end{figure}

\subsection{Dynamic experiments}
\label{sec.expe.dyn}

    \paragraph{Memory-constrained cluster}
    The makespan in case of no recomputation becomes invalid as soon as at least one task finds itself in an invalid memory size
    situation - that is, if the scheduler assumed its memory to be smaller than it actually was and assigned it to a processor
    with not enough memory capacity.
    Due to an extremely constrained memory in this cluster, only 134 experiments out of 1160 for all algorithm variants
    ended with a valid makespan without recomputation.
    In case of \heft, $14$ valid initial makespans were computed.
    Out of them, $13$ managed to stay valid until the end, after all the update requests.
    The same $13$ experiments ended with a valid makespan in case of no recomputation.
    So, these workflows required so few resources that there was no point in re-scheduling them.
    %$13$ experiments ended with a valid makespan after recomputation in case of \heft, which corresponds to $4.4\%$ success
    %rate. However, there were only $14$ successful \heft experiments overall in this setup, so for the successful experiments
    %only, the success rate of not recomputing is $92\%$.
    In case of \heftmm, all $290$ workflows could be scheduled initially and all the schedules remained valid until the end.
    $16$ experiments ended with a successful makespan without recomputation.
    The rate of successful schedules without recomputation is therefore~$5.5\%$.
    \heftblc could produce $142$ valid initial schedules and kept $141$ of them valid until the end.
    $50$ experiments were successful without recomputation, too, making it $35\%$ out of all successful final schedules.
    \heftbl kept $105$ schedules valid until the end out of $110$ initial valid schedules.
    $55$ were able to be kept valid until the end even without recomputation, roughly $50\%$ of all successes.

    \heftbl and \heftblc were successful on smaller workflows and failed on larger ones, the success of the strategy without
    recomputation was limited to even the smallest of these smaller workflows.
    So, the strategy without recomputation delivered valid makespans (independently of the algorithm) on $56$ original
    workflows with $<100$ tasks, $47$ 200-task ones, $25$ 1000-task ones and $6$ 2000-task ones.

    Figure~\ref{fig:updates-ms} shows the increase in makespan in case of no recomputation for these experiments.
    With growing size of the workflow, the excess makespan of not recomputing grows - from $13.9\%$ to $20\%$ for \heftbl,
    from $12.7\%$ to $18.7\%$ on \heftblc (but there is no data for the 2000-task workflows in this case),
    $12.1\%$ to $23.5\%$ for \heftmm.
    The larger variations for \heft can be explained by the small amount of data -- for instance, there were only 2 200-task
    workflows for this case.

    \begin{figure}[tb]
        \centering
        \includegraphics[width=0.8\columnwidth] {images/UpdatesMss2}
        \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm. Smaller is better.}
        \label{fig:updates-ms}
        \vspace{-0.3cm}
    \end{figure}

\subsection{Running times of the heuristics}
\label{sec.expe.t}

    To be able to answer the runtime system without holding it up for too long, the scheduler needs to 
    be able to compute a schedule rapidly. %provide a fast answer.
    The bottom-level-based heuristics \heftbl and \heftblc provide smaller running times than \heftmm, 
    and also scale better with growing workflow
    sizes~(see Fig.~\ref{fig:runtimes-log}).
    Their running times are similar and grow from tens of milliseconds for the smallest workflows to $25-27$ seconds 
    for the largest workflows.
    \heftmm, however, needs to compute a memory-optimal traversal of the entire workflow to compute the ranks 
    of the tasks,
    so its running time scales from also tens of milliseconds for the smallest workflows, 
    to thousands of seconds - $1172.7$ for \numprint{20000}-task workflows and
    $2994.9$ seconds for the largest, \numprint{30000}-task workflows.
    This increased running time is, though, offset by the unique $100\%$ success rate this algorithm 
    obtained when scheduling large
    workflows in difficult (memory-constrained) setups.

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/runtimes-logarithmic}
        \caption{Running times of the heuristics. The y axis is logarithmic.}
        \label{fig:runtimes-log}
        \vspace{-0.3cm}
    \end{figure}

    %\begin{figure}[tb]
    %    \centering
    %    \includegraphics[width=1.1\columnwidth] {images/runtimes-absolute}
    %    \caption{Running times of the heuristics and the baseline.}
    %    \label{fig:runtimes-abs}
    %    \vspace{-0.3cm}
    %\end{figure}

%    \subsubsection{Summary}

\section{Conclusion} %  \skug{Full: 0, Polished: 0}}
\label{sec:conc}

We have formalized a scheduling problem in memory-constrained environments, where tasks
may exceed the memory available on a processor and resort to communication buffers
to store and communicate data between processors. In this context, we have 
designed three memory-aware HEFT-based heuristics, which account for memory
constraints while scheduling tasks.

Two heuristics rely on an ordering of tasks based on the bottom level of tasks, with the objective
of minimizing the makespan and hence scheduling critical tasks first. The third one, \heftmm,
goes one step further and handles tasks in an order dictated by an efficient traversal
of the graph in terms of memory requirements, hence reducing the memory used
by the schedule. Experimental results on a large set of workflows coming from real-life applications
demonstrate that the memory-aware heuristics are  successfully producing valid schedules. 
In the most memory-constrained setting, \heftmm succeeds to schedule even the largest workflows,
while other heuristics return invalid schedules exceeding the memory capacity of processors. 
It comes at a price of a makespan that is not quite as good as the one obtained by \heftbl
and \heftblc, which focus on makespan by ordering tasks by bottom level. 
As expected, the baseline \heft that is not memory aware returns invalid schedules in almost
all settings, except for very small workflows.

%Hence, \heftbl and \heftblc can schedule all workflows in a normal cluster and approximately half of them in a memory-constrained
%    one.
%    They provide a running time that scales well with the growing size of a workflow.
%
%    \heftmm can schedule even the largest workflows even on the memory-constrained cluster.
%    It is the only heuristic that has $100\%$ success rate in our experiments.
%    However, the makespans it produces are larger than those of the the other two.

Another key contribution is that we have adapted these heuristics for a dynamic setting, where exact
task parameters (execution times and memory requirements) are not know in advance. 
We have implemented a runtime system that interacts with the scheduler, returning exact
parameter values once a task arrives in the system, while only estimates are known for future tasks. 
Some preliminary experiments have been conducted in this setting, and demonstrated that
it is necessary to adapt the schedule on the fly in order to avoid an execution failure because
of a shortage in memory. To the best of our knowledge, this is the first study of adaptive
algorithms accounting for memory constraints.

This work could be extended in several directions. First, the model could be refined to include
heterogeneous bandwidths, while we consider a homogeneous communication network. 
More importantly, it would be interesting to consider other types of variability, for instance
if new tasks appear in the graph (or disappear), or if there is variability in the platform, 
with processors arriving and departing. We believe that we could adapt the current approach
by recomputing schedules on the fly, and we plan to perform a new set of experiments
to further assess the impact of dynamic scheduling.

%
%\begin{itemize}
%  \item Extend to heterogeneous bandwidths. 
%  \item Further variability commented out from the paper: changing DAG structure; changing platform...
%\end{itemize}

\balance
    \bibliographystyle{abbrv}%IEEEtran}
    \bibliography{references}

\end{document}