### 2025-03-07 main-disk-model.tex
main-disk-model.tex cleaned until II.C

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{algorithmicx}
%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{numprint}
\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{balance}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\include{macros}

\newcommand{\algo}[1]{\textsc{#1}}
\newcommand{\bottomlevel}[1]{\underline{l}_{#1}} % underline short italic
\newcommand{\criticalpath}{\mathcal{P}}
\newcommand{\parents}[1]{\,\Pi_{#1}}
\newcommand{\children}[1]{\,C_{#1}}
\newcommand{\cluster}{\,\mathcal{S}}

\newcommand{\heft}{\algo{HEFT}\xspace}
\newcommand{\heftmm}{\algo{HEFTM-MM}\xspace}
\newcommand{\heftbl}{\algo{HEFTM-BL}\xspace}
\newcommand{\heftblc}{\algo{HEFTM-BLC}\xspace}

\newcommand{\MM}{M}
\newcommand{\MC}{MC}
\newcommand{\rt}{rt}
\newcommand{\curM}{curM}
\newcommand{\curC}{curC}
\newcommand{\PD}{PD}

\newcommand{\bw}{bw}
\newcommand{\br}{br}
\newcommand{\Moff}[1]{m^{\text{off}}_{#1}}

\newcommand{\skug}[1]{{\color{blue}[SK: #1]}}
\newcommand{\hmey}[1]{{\color{red}[HM: #1]}}
\newcommand{\AB}[1]{{\color{purple}[AB: #1]}}
\newcommand{\willchange}[1]{{\color{orange}[AB: #1]}}

\renewcommand{\iec}{i.e., }
\begin{document}

    \title{Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\
%{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
%should not be used}
    % \thanks{Identify applicable funding agency here. If none, delete this.}
    }

%\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID} }

    \maketitle

    \begin{abstract}

      TODO

    \end{abstract}

%    \begin{IEEEkeywords}
%        DAG, Heterogeneous platform, Adaptive scheduling, Memory constraint.
%    \end{IEEEkeywords}

\section{Introduction}

    %%% CONTEXT %%%

    %%% MOTIVATION %%%

    %%% CONTRIBUTION %%%
    %\paragraph*{Contribution} 
    The main contributions of this paper are both algorithmic and experimental:
    \begin{itemize}
    \item We formalize the problem with memory constraints on heterogeneous processors, where a disk space
   stores files evicted from memory that will be later used by another processor.
    \item We design static and dynamic (adaptive?) heuristics
    \item We implement ...
    \item We perform extensive simulations,
    \item In the dynamic setting, we
    \end{itemize}

Todo: list what we do in each chapter

\section{Related work}
    \label{sec:related-work}

   Todo

\section{Model}
\label{sec:model}

\subsection{Workflow}
\label{sec.mod.work}

    Todo: this section holds. Needs rewriting to avoid plagiarism accusatons. Also insert other references from related work

Our workflow is a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
$E$ is a set of directed edges of the form $e=(u,v)$, with \mbox{$u,v\in V$}, expressing precedence constraints between tasks.
Each task~$u \in V$  is performing $w_u$ operations, and it also
requires some amount of memory to be executed, denoted as~$m_u$.
Each edge $e=(u,v) \in E$ has a cost~$c_{u,v}$ that corresponds to the size of the output file written by task~$u$ and used as input by task~$v$.

Note that $m_u$ is the total memory usage
of a task during its execution, including input and output files currently being read and written,
and hence it is greater than the total size of input files
(received from the predecessor tasks) and than
the total size of  output files (sent to successor tasks):
$$   m_u \geq \max \left\{ \sum_{v:(v,u)\in E}c_{v,u}, \sum_{v:(u,v)\in E} c_{u,v} \right\} . $$

The predecessors of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of predecessors is
$ \parents{u} = \{v \in V: (v,u) \in E\}$. A task without predecessors is called a {\it source task}.
The successors of a task~$u$ are the tasks following~$u$ directly according to the precedence constraints, i.e.,
$ \children{u} = \{v \in V: (u,v) \in E\}$. A task without successors is called a {\it target task}.
Each task may have multiple predecessors and successors.

Furthermore, we place ourselves in a context where we do not have perfect knowledge
of the length of the tasks, i.e., $w_u$,
before the tasks complete their execution,
but only estimates~\cite{rahman2013,GARG2015256}.

Hence, scheduling decisions are made on these estimated parameters, and
may be reconsidered at runtime when a task completes its execution.

    \subsection{Execution environment}
    \label{sec.mod.plat}

    Todo: this section holds and is new (in parts).

The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
consists of $k$ processors $p_1, \dots, p_k$.
Each processor $p_j$ ($1 \leq j \leq k$) has an individual memory of size $M_j$ and a speed~$s_j$.
All processors have access to a shared disk of unlimited size, where they can write
and read data: Processor~$p_j$ has a bandwidth $\bw_j$ (resp.~$\br_j$) to write  (resp. read),
and hence it takes a time $\frac{c_{u,v}}{bw_j}$ to write the data produced by task~$u$
on~$p_j$ for task~$v$, 
if this data is evicted from memory.

All communications happen over the disk -- after one processor has finished writing data there, another one can read
it to load the data on its individual memory. 
Hence, we can decide to evict some data from the individual processor memories if it has been 
written onto the disk, 
in order to free some space in the individual memories.
This data can later be read back into memory of the same or any other processor at any time.

If the entire memory required by  task ~$u\in V$ fits in the available memory of processor~$p_j$,
    then the execution time of this task on this processor is expressed as $\frac{w_u}{s_j}$.
However, if the memory requirement of the task exceeds the available memory on the processor, 
then task can still be executed there, but it will be slowed down. Indeed,

the part of memory requirement that exceeds the available memory on processor~$j$ for task~$u$, 
denoted by $\Moff{u}$,
will be offloaded to the disk, slowing down the execution:

    \[
        w_{real} = \frac{w_u}{s_j} +  q_j \times  \Moff{u},
    \]
where $q_j$ is the slowdown coefficient for $p_j$.

Multiple processors can communicate with the disk in parallel, but the communication of each processor  with the disk is sequential.
The processor can read and write at the same time, but only one file can be read and only one written at any point in time.
We keep track of the current ready times of each processor $j$, ready time for computation $\rt_j^c$ (when one task has finished
computing and another one can start),
ready time for writing $\rt_j^w$ (when a file has been written to disk and another file can be written), and ready time for
 reading~$\rt_j^r$.
Initially, all the ready times are set to~$0$.
We also keep track of the currently available memory, $availM_j$ on the processor memory.

Furthermore, $\PD_j$ is a priority queue with the {\em pending data}
that are in the memory of processor~$p_j$,  but that may be evicted to the disk
if more memory is needed on~$p_j$.
They are ordered by non-decreasing size and correspond to some files~$c_{u,v}$.
When scheduling a task, the decision has to be made to execute this task on a processor with enough memory
without slowdown, but with possibly more time to read the input files, or to execute the same task
on a processor with less memory (and hence a slowdown), but potentially with a higher processor speed 
or with less times spent reading files.

\textcolor{orange}{ TODO: Move the next paragraph to Algorithms.} We use the \algo{memDag} algorithm developed by Kayaaslan \etal~\cite{KAYAASLAN20181} to compute
the memory requirement; it transforms the workflow into a series-parallel graph
and then finds the traversal that leads to the minimum memory consumption.
\AB{When do we need this? Probably more in Algorithms, right? }

    \begin{table}
        \begin{center}
            \begin{tabular}{rl}
                \hline
                \textbf{Symbol}                       & \textbf{Meaning}                                         \\
                \hline
                $G = (V, E)$                          & Workflow graph, set of vertices (tasks) and edges        \\
                $\parents{u}$, $\children{u}$         & Predecessors of a task $u$, successors of a task $u$            \\
                $m_u$                                 & Memory weight of task $u$                                \\
                $w_u$                                 & Workload of task $u$  (normalized execution time)          \\
                $c_{u,v}$                             & Communication volume along the edge $(u,v)\in E$         \\
                $F$, $\mathcal{F}$                    & A partitioning function and the partition it creates     \\
                $V_i$                                 & Block number $i$                                         \\ %\wrt~some $F$   \\
                $\cluster$, $k$                    & Computing system and its number of processors           \\
                $p_j$, proc($V_i$)                          & Processor number $j$, processor of block $V_i$                 \\
                $M_j$, $MC_j$, $s_j$                               & Memory size, comm. buffer size, and speed of proc.\ $p_j$                          \\
                $\beta$                     & Bandwidth in the compute system                                \\
                $\bottomlevel{u}$                      & Bottom weight of task $u$ \\
                $\mu_G$, $\mu_i$ & Makespan of the entire workflow $G$ and of a block $V_i$               \\
                $\Gamma = (\mathcal{V}, \mathcal{E})$                      & Quotient graph, its vertices and its edges        \\
                $r_u$, $r_{V_i}$                            & Memory requirement of task $u$ and of block $V_i$                 \\
                \hline
            \end{tabular}
        \end{center}
        \caption{Notation} \label{tabnotation}
    \end{table}

\subsection{Optimization problem}
\label{sec.mod.pb}

In the {\bf offline setting}, the goal is to find a {\em schedule} of the DAG~$G$ for the $k$ processors,
so that the makespan (total execution time) is minimized. Formally, a schedule contains:
\begin{itemize}
\item for each task, its processor allocation and starting time, 
as well as the amount of data offloaded for the execution of the task; 
\item for each read and write operation to disk, the starting time of the operation; 
\item for each data and processor, the time intervals during which the file is available
on the local memory of the processor (once it has been generated or read, and before 
it is used or evicted). 
\end{itemize}

\AB{I'm confused again about the offloading model, we need to talk}

%\AB{no more memory constraint, now the lack of memory just slows down the execution, correct?
%I'll probably add a few sentences here...}
% while respecting memory constraints. \skug{do we still formulate it like that?}
    %If a processor runs out of memory to execute
%a task mapped on it, the schedule is said to be {\em invalid}.

\medskip
We also consider the {\bf online setting} where tasks are subject to variability, 
and then we know the exact time required to complete a task only when 
it ends its execution. Hence, we aim at minimizing the actual makespan
achieved at the end of the execution, while scheduling decisions 
have to be taken building
on the estimated task parameters.
In this case, we do not build the whole schedule offline, but we build
it on the fly, as tasks become ready (their predecessor tasks have been completed). 
%\AB{Probably need to clarify offline vs online problem at some point...}

Note that the offline problem is already NP-hard even in the homogeneous case and 
without memory constraints (i.e., no need to offload data onto disk), and even for
a graph without precedence constraints (independent tasks). 
%because of the DAG structure of the application. 
Hence, we focus on the design of efficient scheduling heuristics. 
%\hmey{Mention NP-hardness due to being more general than NP-hard problem?}

%    \paragraph{Workflow-related changes}
%
%    \begin{itemize}
%        \item A task $v$ takes longer or shorter to execute than planned: its time weight $w_u$ changes to $w'_u$.
%        \item A task $v$ takes more or less memory to execute than planned: its memory requirement $m_v$ changes to $m'_v$.
%
%    \end{itemize}
%
%    The following changes are not a part of this article's scope:
%
%    \begin{itemize}
%        \item The workflow structure changes: edges or tasks come in or leave.
%    \end{itemize}
%
%    \paragraph{Execution environment-related changes }
%
%
%    \begin{itemize}
%        \item A processor exists the execution environment: $k$ decreases and $\cluster$ changes.
%        \item A processor enters the execution environment: $k$ increases, $\cluster$ gets a new processor with possibly new memory requirement and processor speed.
%
%    \end{itemize}
%
%    The following changes are not a part of this article's scope:
%
%    \begin{itemize}
%        \item Processor characteristics change: the memory requirement or speed become bigger or smaller
%    \end{itemize}
%
%    \subsection{Time of changes }
%
%    We consider discrete time in seconds.
%    The time point(s) at which the changes happen is unambiguously defined.
%
%    For any task $v$, its runtime equals its time weight divided by the speed of the processor $p_j$ it has been assigned to: $w_v/s_j$.
%    The start time of any task $v$ is its top level($\bar{l}_v$), or the difference between the maximum bottom level in the workflow (the makespan of the workflow) and the task's own bottom level: $\bar{l}_v = \mu_\Gamma - \bottomlevel{v}$.
%    The start time of the source task in the workflow is zero.
%    The end time of a task $v$ is its start time and its runtime: $\bar{l}_v + w_v/s_j$
%
%    \subsection{Changes and knowledge horizon - important questions TBA}
%
%    Given a valid mapping of tasks to processors, we can say what we predicted would happen at any given time point $T$: what tasks have been executed, what have not finished or have not even started.
%
%    At the point of change, we know that some tasks that finished took longer than expected ($w_v$ are bigger) or shorter.
%    However, how do we model the following:
%    \begin{itemize}
%        \item Do we know the new weights of currently running tasks and tasks that have not yet started? This means, do we foresee into the future or do we assume that all weights on unfinished tasks remain the same?
%        \item A change in memory requirements can mean that the assignment had been invalid. Do we assume that these tasks failed and we need to rerun them?
%        \item How many times of change do we model - one per workflow run, or multiple?
%        \item At what time does the change and reevaluation happen - is it a fixed (random?) point of time or is it workflow-dependent (say, after 10\% of the workflow is ready)?
%    \end{itemize}

\section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
\label{sec:heuristics}
\subsection{\skug{New scheduling approach - general description}}

    In the first step, we order the tasks by non-increasing ranks (e.g. their bottom levels).
    In the next step, we repeatedly schedule ready tasks.
    For each task, we first choose the best processor to put it on - the one that minimizes the expected finishing time.
    For each processor, we compute this earliest finishing time by first calculating the earliest starting time on this processor.
    It is computed as the maximum of the ready time of computation on this processor, and the time to read input files.
    For the currently reviewed processor, we preliminarily schedule necessary file writes on other processors holding input files,
    and file reads on the current processor.
    Doing so, we respect the ready times on writes on other processors and reads on the current processor.
    We possibly plan ``into the past'' with ready and writes.

    The earliest finish time is calculated by adding the execution time (possibly with delay due to memory overflow) to the start time.
    Further scheduling decisions happen when another task finishes.
    \skug{check this: when a task finishes earlier or later than planned, we review our earlier scheduling decisions and may
    move the task execution to another processor if this proves more beneficial.}

    When a task finishes, we look at all of its sucessors, rather than only the ready ones.
    For tasks that lie further in the future, we try to plan rather early.
    Therefore, we employ lazy writes.
    We try to write the files for these tasks to the disk early.
    So, we preliminarily schedule these file writes. However, they are not high priority and can be cancelled.
    We keep two ready times on writes, one soft, with lazy writes, and another hard onw, with scheduled writes.
    If a task requires a write during the lazy write, we cancel it and move it to later point in time.

    \skug{For each edge, we also keep information on where it is currently stored - in memory of a processor (with processor id),
        on the disk or nowhere - if this file has not yet been generated or has already been consumed by the successor task.}

\subsection{Data structures }
    To optimally represent the timeline of the execution of a workflow, we represent the execution as a series of events
    with attached unique timestamps.
    The events are kept in the priority queue ordered by timestamp.
    Events are created when other events are triggered and are being inserted into the queue.
    When their timestamp is the smallest, they are then triggered, possibly creating new events.
    The execution starts with placing the event of starting the starting node into the queue with the timestamp 0.
    The execution ends when there are no events in the queue.

    An event can happen on finish or start of a task execution or an io action (file read or write).
    Each event has several parameters:
    \begin{itemize}
        \item Event type: OnTaskStart, OnTaskFinish, OnIoStart, OnIoFinish
        \item The workflow entity associated with the event: the task to be executed or an edge whose file needs to be read or written
        \item The processor id on which the action should take place
        \item Expected and actual time of firing.
        \item A set of predecessor events. The event is called ready if all its predecessors have been finished.
        \item A set of successor events to keep track of.
        \item For a write event, a boolean if this file should be removed from memory after being written.
    \end{itemize}

    Per processor,  in addition to keeping the fixed values (memory, speed), we keep track also of the current state in its memory.
    Because during and after execution of each even the state of the processor memory changes, we need to keep snapshots of what files
    were in memory and how much of it was available.
    Because keeping such snapshots for each timestap is too complicated, we only keep two timestamps.
    First, we keep the information of available memory and tasks pending in memory during the execution of the last
    task scheduled to this processor.
    Then, we keep the same information on what it will be immediately after this task finishes.

    \subsubsection{On Task Finish}
    When a task finishes, it first cleans after itself:
    \begin{itemize}
        \item It frees its memory: it increases the available memory on the processor by the amount it occupied
        \item If there had been a memory overflow onto disk, then this memory is being freed implicitly
        \item it sets its status to finished in the workflow.
    \end{itemize}
    It the goes over its successor events (Tasks and IO actions) and removes itself from their predecessor lists.
    It also updates their actual trigger times with its actual finish time.
    Successor tasks can be tasks scheduled on the same processor after this one.
    Successor io events can be reads that could not start until the task finishes and frees its memory.
    \skug{can writes be successor events of a task?}

    Then task goes over its successor tasks in the workflow and schedules ready ones.
    Ready tasks are those whose all predecessors have the finished status.

    For each ready task, the scheduler tries to tentatively put it on each processor in the cluster and chooses the processor
    that promises the earliest finish time.
    The finish time on each processor is calculated as follows:

    \paragraph{Earliest start and finish time}
    We find out when the last task on this processor finishes.
    If there is enough memory to start our task after that, then the end of previous task is the earliest start time for our task.
    If files need to be evicted first, we need to find a balance between evicting tasks (and waiting for their writes to finish),
    then running our task, and starting the task immediately and letting it run longer due to memory overflow.
    We compare the following cases:
    \begin{itemize}
        \item Evict everything, then start task
        \item Greedily evict the largest file, starting the task
        \item Not evicting anything, starting the task
    \end{itemize}
    Whatever combination gives the best estimated finishing time for the task, we note the estimated starting time of the task in this case.
     The finish time is the earliest start time  plus computation time of the task.

    \paragraph{Incoming edges} For each incoming edge, its location is determined.
    If the edge is in the memory of the target processor, nothing needs to be done.
    The estimated start and finish times for the task are unaffected.

    If the edge is on the disk, it first needs to be read into the memory of the target processor.
    We delay the reads to avoid holding unnecessary files in the memory.
    When we preliminarily schedule a read, we first assess its length and other actions that happen on this processor in the meantime.
    So, first estimated start of the read is \skug{the maximum between the desired start time (estimated start time of the task minus the estimated length of the read),
    and the feasible one (when the lastest read event on this processor finishes)}.

    We then see if this start of the read is happening during the runtime of the previous task or already after it (during the writes).
    If the starting time of the read is during the execution of the previous task and there is enough memory to read this file, then we move the read
    forward to the beginning of this previous task.
    If there is not enough memory, then the entire read is rescheduled for after the task finishes.
    The read will then be bottleneck and affect the eraliest start time of the desired task.

     If the edge is in another processor's memory, it needs to be first written to disk, then read by the target processor.
    So we look at the latest write on that processor, preliminarily schedule the write to disk and then proceed like in the
    previous case, reading the file from disk.

   \paragraph{When processor is chosen} When a processor is chosen, the scheduler inserts new events into the queue:
    \begin{itemize}
        \item For the task, OnTaskStart and OnTaskFinish events with the estimated trigger times, the OnTaskFinish event being dependent on the OnTaskStart event.
        \item For each io operation necessary for the execution: OnIoStart event of the corresponding type and on corresponding processor
        \item It connects all the event with predecessor-successor relations.
        \item If files needed to be evicted
    \end{itemize}
    The scheduler also updates the ready times of the processors and updates the new currently last events there.

    \subsubsection{On Task Start}
    When a task start event is triggered, it first determines if it can start.
    If some predecessor tasks have not yet finished (the predecessor set is not empty), then the actual time of this event
    is set to the maximum of the estimated end times of predecessor events and this event is reinserted in the queue
    with this new time.
    \skug{We  also update the actual times of all its successor events.}

    If the task can start, the scheduler determines the deviation of the runtime of the task.
    It applies the deviation function on its estimated runtime (estimated finish minus estimated start time).
    It then adds this deviated runtime to its actual start time and receives an updated actual finish time.
    With this time, it updates the OnTaskFinish event of this task.

    \subsubsection{OnIoFinish}
    When the io operation finishes, it removes itself from the predecessor sets of all its successor events.
    It also sets their actual time to its own actual finish time.

    \subsubsection{OnIOStart}
    When the IO operation starts, it updates the corresponding IoFinish event

    After being triggered and completed, each event is deleted from the queue.

--------------------------------------

%    The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
%    but favor processor reuse as is done in HEFT.
%    Furthermore, this would allow us to handle variability on the fly, by updating
%    the bottom levels if some parameters vary, and computing the schedule
%    only for the near future...

%In order to be able to easily adapt to variability of task parameters, 
We design 
variants of HEFT that account for memory usage and aim at minimizing the makespan.
First, we present in Section~\ref{sec.heft} the baseline HEFT heuristic that does not account for the memory
(and hence, may return invalid schedules that will not be able to run on the platform
by running out of memory).  Then, Section~\ref{sec.heftm} focuses on the presentation of the novel
heuristics, including eviction strategies to move some data in communication buffers
in case there is not enough memory available on some processors.

    \subsection{Baseline: original HEFT without memories}
\label{sec.heft}

    Original HEFT does not consider memory sizes.
    The solutions it provides can be invalid if it schedules tasks to processors without sufficient memories.
    However, these solutions can be viewed as a ``lower bound'' for an actual solution that considers memory sizes.

    HEFT works in two stages.
    In the first stage, it computes the ranks of tasks by computing their non-increasing bottom levels.
    The bottom level of a task is defined as
    $$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
    (the max is 0 if there is no outgoing edge).
    The tasks are sorted by non-decreasing ranks.

    In the second stage, the algorithm iterates over the ranks and tries to assign the task to the processor where it
    has the earliest finish time.
    We tentatively assign each task~$v$ to each processor~$j$.
    The task's starting time $st_v$ is dictated by the maximum between the ready time of the processor~$rt_j$,
    and all communications that
    must be orchestrated from predecessor tasks $u\notin T(p_j)$.
    The starting time is then:

 {\footnotesize{   \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v)}\{ FT(u)+ c_{u,v} / \beta , rt_{proc(u), p_j} + c_{u,v} / \beta  \} \} } \]}}

    Finally, its finish time on $p_j$ is
    $FT(v,p_j) = st_v + \frac{w_v}{s_j}$.

    Once we have computed all finish times for task~$v$,
    we keep the minimum $FT(v,p_j)$ and assign task~$v$
    to processor~$p_j$.

    \textit{Assignment to processor. }
    When assigning the task, we set the ready time $rt_j$ of  processor~$j$ to be the finish time of the task.
    For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
    communication buffers $rt_{j', j}$ for every predecessor $u$'s processor $j'$: we increase them by the
    communication time $c( u,v) / \beta$.

    \subsection{Memory-aware heuristics}
    \label{sec.heftm}
    Like the original HEFT, the memory-aware versions of HEFT consist of two stages:
    first, they compute the task ranks,
    and second, they assign tasks to processors in the order defined in the first stage.
    We consider three variants of HEFT accounting for memory usage (HEFTM), which only
    differ in the order they consider tasks to be scheduled in the first stage.

\smallskip
    \noindent{\bf Compute task ranks. }

We design three variants of memory-aware HEFT:
\begin{itemize}
\item    HEFTM-BL orders tasks by non-increasing bottom levels, where the bottom
    level is defined as
    $$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
    (the max is 0 if there is no outgoing edge).

\item    HEFTM-BLC %: from the study of the fork (see below), it seems important
   %  to also account for the size of the data as input of a task,
    is  giving more priority at tasks with potential large incoming communications,
    hence aiming at clearing the memory used by files as soon as possible,
    to have more free memory for remaining tasks to be executed on the processor.
    Therefore, for each task, we compute a modified bottom level accounting for communications:
    $$blc(u) = w_u + \max_{(u,w)\in E} \{c_{u,w} + blc(w)\} + \max_{(v,u)\in E} c_{v,u}   . $$

%    \skug{avoid having mixed ranks, when the memory size of the lower task is not taken into account}

\item   Finally, HEFTM-MM orders tasks  in the order returned by %as dictated by MinMem.
the  \algo{memDag} algorithm~\cite{KAYAASLAN20181}, which corresponds to a traversal
of the graph that minimizes the peak memory usage.

\end{itemize}

\smallskip
  \noindent  {\bf Task assignment. }

    Then, the idea is to pick the next free task in the given order,
    and greedily assign it to a processor, by trying all possible options
    and keeping the most promising one. We first detail how a task
    is tentatively assigned on a processor, by carefully accounting for the memory usage.
    Next, we explain the steps to be taken to effectively assign a task on a given processor.

    \medskip
    \noindent{\em Tentative assignment of task~$v$ on $p_j$.}\\
    {\bf Step 1.} First, we need to check that for all predecessors~$u$ of~$v$ that are mapped
    on~$p_j$, the data $c_{u,v}$ is still in the memory of~$p_j$,
    i.e., $c_{u,v}\in PD_j$. Otherwise, the finish time is set to~$+\infty$ (invalid choice).

    \smallskip
    \noindent{\bf Step 2.} Next, we check the memory constraint on~$p_j$, by computing
    $$Res = availM_j - m_v - \!\!\!\! \sum_{u \in \Pi(v), u\notin T(p_j)}  \!\!\!\!\!\!\!\!\{c_{u,v}\}
    - \sum_{w\in Succ(v)} \!\!\!\!\!\! \{c_{v,w}\}.$$

    $T(p_j)$ is the set of tasks already scheduled on $p_j$; by step 1, their files are
    already in the memory of~$p_j$. However, the files from the
    other predecessor tasks must be loaded in memory before executing task~$v$,
    as well as $m_v$ and the data generated for all successor tasks.
    $Res$ is then checking whether there was enough memory; if it is negative,
    it means that we have exceeded the memory of~$p_j$ with this tentative
    assignment.

    In this case ($Res <0$), we try evicting
    some data from memory so that we have enough memory to execute task~$v$.
    We need to evict at least $Res$ data.
    We propose a greedy approach, evicting the largest files of $\PD_j$ until $Res$ data has been evicted. 
    A variant where the smallest files are evicted first has been tested, and it led to comparable results. 
%    
%    in order to avoid costly communications.
%    \AB{FYI We initially discussed evicting the largest files, but this leads to
%    large communications and does not seem efficient after all... Maybe we can think of another
%    approach that would take into account both data size and bottom level...}
    While tentatively evicting files, we remove them from the list of pending memories and move them into a list
    of memories pending in the communication buffer.
    We keep track of the available buffer size, too -- each time a file gets moved into the pending in buffer, 
    the available buffer size is reduced by its weight.

    If we still do not have enough memory after having tentatively evicted all files from $\PD_j$,
    or if while doing so we exceeded the size of the available buffer,
    we set the finish time to~$+\infty$ (invalid choice).

    \smallskip
    \noindent{\bf Step 3.} We tentatively assign task~$v$ on $p_j$.
    Its starting time $st_v$ is dictated by the maximum between $rt_j$, and all communications that
    must be orchestrated from predecessor tasks $u\notin T(p_j)$.
    The starting time is therefore:\\[-.7cm]

   {\footnotesize{ \[ST(v, p_j) = \max{ \{rt_j, \max_{ u \in \Pi(v), u\notin T(p_j)}\{ FT(u) , rt_{proc(u), p_j}\} + c_{u,v} / \beta \} .} \]}}

  \noindent  Finally, its finish time on $p_j$ is 
    $FT(v,p_j) = ST(v, p_j) + \frac{w_v}{s_j}$.

    \medskip
    \noindent{\em Assignment of task~$v$.}\\
    Once we have computed all finish times for task~$v$,
    we keep the minimum $FT(v,p_j)$ and assign task~$v$
    to processor~$p_j$.
    In detail, we:
    \begin{itemize}
        \item  Evict the file memories that correspond to edge weights that need to be evicted to free the memory.
        We remove these files from pending memories
        $PD_j$, add them to pending data in the communication buffer, and reduce the available buffer size accordingly.
        \item    Calculate the new $availM_j$ on the processor.
        We subtract the weights of all incoming files from predecessors assigned to the same processor,
        and add the weights of outgoing files generated by the currently assigned task.
        \item  For every predecessor of~$v$ that has been assigned to another processor, we adjust ready times on
        communication buffers $rt_{j', j}$ for the processor~$j'$that the predecessor $u$ has been assigned to: we increase them by the
        communication time $c( u,v) / \beta$.
        We also remove the incoming files from either the pending memories or pending data in buffers of these other
        processors, and increase the available memories or available buffer sizes on these processors.
        \item We compute the correct amount of available memory for~$p_j$ (for when the task is done).
        Then, for each predecessor that is mapped to the same processor, 
        we remove the pending memory corresponding to the weight of
        the incoming edge, also freeing the same amount of available memory (increasing $availM_j$).
        For each successor, we rather add the edge weights to pending memories and reduce $availM_j$ 
        by the corresponding amount.
    \end{itemize}

\hmey{BEGIN FORK}
    \subsection{The fork}
    We look at the behavior of these heuristics on a fork graph,
    where there is a root task~$T_0$, producing $n$ files $f_1, \ldots, f_n$
    to be used by tasks $T_1, \ldots, T_n$ ($f_i = c_{0,i}$).

    Without memory, this problem is NP-complete; this is equivalent
    to 2-partition if the tasks have $w_i=a_i$, and all files are of size~$f_i=0$,
    and with two processors. Half of the tasks must be sent to the processor
    on which $T_0$ is not executed, and the optimal makespan is
    $w_0+\frac{1}{2}\sum_{1\leq i \leq n} w_i$.

    However, with an infinite number of identical processors, it can be
    solved in polynomial time: sort tasks by non-decreasing $f_i+w_i$;
    the $k$ tasks with smallest $f_i+w_i$ are then sent to another processor,
    while the remaining $n-k$ tasks are executed locally (try all values of $k$).

    With heterogeneous processors, it is probably NP-complete again
    because we could ensure that there are only two processors fast enough
    and get back to the 2-partition...

    We also had an example where evicting large files first in step 2
    can lead to arbitrarily bad makespan. Consider a fork with $n=2$,
    $f_1=1$, $w_1=2$, $f_2=100$, $w_2=1$, and memory constraint
    imposes that we free one unit of memory before executing one
    of the tasks\ldots Actually the new version with BLC would start
    considering $T_2$ and be fine in this case\ldots

\begin{figure}[tb]
  \centering
  \includegraphics[width=0.6\columnwidth]{images/bad-blc-instance.png} \\[1ex]
  \includegraphics[width=0.999\columnwidth]{images/blc-bad-cost.png} \\[1ex]
  \includegraphics[width=0.67\columnwidth]{images/opt-cost.png}
  \caption{Top: Example instance $I$, middle: BLC applied to $I$ leads to makespan $w_0 + nk/2 + 1+\epsilon$, bottom: optimal solution has makespan $w_0 + n/2 + k$.
  \hmey{TODO: adapt resolution of figure}}
  \label{fig:fork-bad-blc-instance}
\end{figure}

Figure~\ref{fig:fork-bad-blc-instance} shows an instance where \heftblc\ performs very poorly.
Since the sorting criterion $w_i + c_{0i}$ yields an arbitrary ordering of the tasks $T_i$ (with $1 \leq i \leq n$), we may
assume that the algorithm processes the nodes in level $1$ of instance $I$ from left to right. 
As shown in the middle part of the figure, $T_1$ is processed on $P_1$ due the earliest finishing time on this processor.
$T_2$ is then sent to $P_2$ because it finishes there $\epsilon$ timesteps earlier than on $P_1$. The same rationale applies
to the other odd/even tasks. This leads to a makespan of $w_0 + nk/2 + 1 + \epsilon$. The optimal solution, in turn, as shown in the bottom,
requires only $w_0 + n/2 + k$ timesteps (assuming $k > n\epsilon/2$). This means that \heftblc\ may yield solutions whose 
gap of optimality is $\Omega(k)$.

\AB{Can we prove that we have (maybe) a 2-approximation,
        at least for the fork? What worst-case can we think of? }
\hmey{END FORK}

Here is the new fork example, that would lead to both \heftbl and \heftblc to get arbitrarily bad result. 
Consider a fork with $n=2$, $w_0=0$, $w_1=1$, $w_2=2$. Furthermore, 
$f_1=1$ and $f_2=K$, where $K$ is an arbitrarily large number. 
The memory requirements of tasks are $m_0=K+1$, $m_1=1$, and $m_2=2K$, 
while the memory size on each processor is~$2K$.

Both heuristics would first try to schedule task~$T_2$. At the end of $T_0$ on processor~1,
the memory usage is $1+K$; to execute~$T_2$ on $p_1$, we would go up to $1+2K = f_1+m_2$, which
exceeds the available memory on~$p_1$. If the offloading penalty $q_1$ is very high,
we would rather schedule $T_2$ on $p_2$ at time $2K$ (write $f_2$ on disk, read it back from disk,
assuming write/read bandwidths to be~$1$), hence completing execution at time~$2K+2$. 
Then, $T_1$ is scheduled on $p_1$ and completes at time~$1$. 
\AB{Actually, we can consider having a slow bandwidth as well, but offloading is even worth,
hence that would be enough to take the decision to execute $T_2$ on~$p_2$. }

However, if we schedule~$T_1$ first, we execute it on~$p_1$, free the memory of $f_1$, 
and then we can execute~$T_2$ on $p_1$ as well without offloading, 
hence an optimal makespan of~$3$.

However, \heftmm would behave nicely here, since the traversal to minimize peak memory
would first consider~$T_1$ and lead to the optimal solution.

\section{Dynamic scenario}
\label{sec:dyn}

    In a workflow execution environment, the scheduling method interacts with the runtime environment, which provides information such as resource estimates.
    This information may include memory usage, runtime, graph structures, or the status of the underlying infrastructure.
    In order to ensure that the information is up to date, a monitoring system observes the workflow execution and collects metrics for tasks and the underlying infrastructure.
    By incorporating dynamic monitoring values, e.g., the resources a task consumed, the runtime environment can incorporate the data into the prediction model to provide more accurate resource predictions.
    Also the underlying infrastructure can change during the workflow execution.
    Examples are processor failures, node recoveries, or acquisition of new nodes.
    However, also when the hardware of the infrastructure does not change, the set of nodes provided as a scheduling target might change due to release or occupation in shared cluster infrastructures.
    As infrastructure information and resource predictions are dynamically updated and provided to the scheduler during the workflow runtime, the previous schedule becomes invalid and a new one must be calculated.

    For state-of-the-art memory prediction methods, a cold-start median prediction error for heterogeneous infrastructures
    of approximately 15\% is shown~\cite{malik2013execution}.
    Online prediction methods were able to significantly reduce the error during runtime, with the reduction reaching up to one-third of the cold-start error~\cite{baderDiedrichDynamic2023,witt2019learning}.
%For instance, Nadeen~et~al.\cite{} report an error of 10\%, 11\%, and 15\% while the task prediction errors shows a normal and exponential distribution.
%Bader~et~al.~ report a prediction error between 13\% and 17\% for their method, showing an exponential task error distribution.
% @Svetlana, willst du sowas für deine Experimente? Also die Daten, welche du dann konfigurieren kannst?
    Such a dynamic execution environment requires a dynamic scheduling method where the schedule can be recomputed during the workflow execution.

    \subsection*{Retracing the effects of change on an existing schedule}
    After the monitoring system has reported changes, we need to assess their impact on the existing schedule.
    These changes can invalidate the schedule (\eg if there is not enough memory for some tasks to execute anymore),
    they can lead to a later finishing time (\eg if some tasks are longer and they delay other tasks), or they can have no effect (\eg if new processors
    joined the cluster, but the old schedule did not account for them).
    To assess the impact, we need to retrace the schedule.

    First, we find out if at least one processor that had assigned tasks has exited - this instantly invalidates the
    entire schedule.

    We then iterate over all tasks of the workflow in a topological order - any of the orderings given by rankings BL, BLC or MM
    is a topological ordering.
    We then repeat steps similar to those we did during tentative assignment in the heuristics, 
    except that we do not choose a processor
    anymore, but rather we check whether the current processor assigned to the task still fits.

    For each task $v$, we first assess its current memory constraint $Res$ using Step 2 from the heuristic.
    The factors that affect $Res$ are possible changes in $m_v$, in $c_{u,v}$ from predecessors $u$ 
    or $c_{v,w}$ from successors $w$,
    available memory $availM_j$ on the processor (due to either changed $M_j$ or changed memory requirements 
    from other tasks).
    If originally, $Res$ was positive (no files were evicted from memory into the communication buffer), 
    then it has to stay this way --
    otherwise, evicted files can invalidate next tasks.
    If original $Res$ was negative, then we need to make sure that evicted files still fit into the communication buffer.
    If either $Res$ is newly negative, or the communication buffer is not large enough, this invalidates the schedule.
    We update the $availM_j$ and $availMC_j$ values according to the new memory constraints.

    Then, we can re-calculate the finish time of the task on its processor like in Step 3.
    The factors that affect it are changes in own execution time $w_v$ of the tasks, changed ready time of the processor
    (after delayed previous tasks), and changed communication buffer availability.

    Finally, after having updated the processor's values, we move on to the next task.

%    \subsection{Approximation}
%    \hmey{Rough notes:}
%    Let's use a fork to see how the algorithm behaves and if it provides some approximation. Our current intuition is that, if the memory constraint is ignored, HEFTM-BLc provides a $2$-approximation (to be proved).

\section{Experimental evaluation} %: \skug{Full: 3, Polished: 3}}
\label{sec:expe}

We first describe the experimental setup in Section~\ref{sec:setup}. 
Then, we report results on static experiments to assess the performance
of the memory-aware heuristics in Section~\ref{sec.expe.static}, before
discussing the heuristics behavior in a dynamic setting in Section~\ref{sec.expe.dyn}. 
Finally, we report running times of the heuristics in Section~\ref{sec.expe.t}.

\subsection{Experimental setup} %: \skug{Full: 5, Polished: 4}}
    \label{sec:setup}

    All algorithms are implemented in C++ and compiled with g++ (v.8.5.0).
    The experiments are managed by simexpal~\cite{DBLP:journals/algorithms/AngrimanGLMNPT19} and executed on workstations with 192 GB RAM and 2x 12-Core Intel Xeon 6126 @3.2 GHz
    and CentOS 8 as OS.
    Code, input data, and experiment scripts are available to allow reproducibility of results, at~\url{https://zenodo.org/records/13919214}~and~\url{https://zenodo.org/records/13919302}.

    We first describe the set of workflows used for the evaluation, and then the clusters on which the
    workflows are scheduled.

    \subsubsection{Workflow instances}
  %  \skug{Todo: describe the workflows the same asin the abstract}
    We run  experiments on a dataset that consists of real-world workflows: 
    we use workflows coming from~\cite{lotaru} (atacseq, bacass, chipseq,
    eager and methylseq), and also we increase their size using the WFGen generator~\cite{COLEMAN202216}.

    \paragraph{Workflow graphs}
    For the five real-world workflows, their nextflow definition (see~\cite{ewels2020nf}) was downloaded from the
    respective repository and transformed into .dot format using the nextflow option ``-with-dag''.
    The resulting DAG contains many pseudo-tasks that are only internal representations in nextflow
    (and not actual tasks); that is why we removed them.

    For the size-increased workflows, the graph is produced by the WFGen generator, based on a {\em model workflow} and
    the desired number of tasks.
    We used the real-world workflows as models, except for bacass that lead to errors in the generator.
%
    As number of tasks, we use: 200, \numprint{1000}, \numprint{2000}, \numprint{4000}, \numprint{8000}, \numprint{10000},
    \numprint{15000}, \numprint{18000}, \numprint{20000}, \numprint{25000}, \numprint{30000}.
    We divide the workflows into four groups by size: tiny ones with up to \numprint{200} tasks, small ones with \numprint{1000} to \numprint{8000} tasks,
    middle ones with \numprint{10000} to \numprint{18000} tasks, and big ones with \numprint{20000} to \numprint{30000} tasks.

    \paragraph{Task and edge weights}
    For the real-world workflows, we use historical data files provided by Bader~\etal~\cite{lotaru}.
    The columns in these files are measured Linux PS stats, acquired during an execution of a nextflow workflow.
    Each row corresponds to an execution of one task on one cluster node.
    Since the operating system cannot distinguish between (a) the RAM the task uses for itself and (b) the RAM it uses
    to store files that were sent or received from other tasks, the values in the historical data are total memory requirements (input/output files plus memory consumption of the computation).
    In a similar manner, the historical data provided by~\cite{lotaru} do not store the actual weights of edges between tasks, but only the overall
    size of files that the task sends to all its children.

    For each task, historical data can contain multiple values, obtained from the runs with different input sizes.
    The same workflow can require different memory capacity and take different time to execute
    depending on the size of its input.
    We simulate these various runs by obtaining values corresponding to each input size.
    For each of the four families, there are five input sizes, so we run each workflow in five variants corresponding to these inputs

    Not all tasks have historical runtime data stored in the tables.
    In fact, for two workflows, Bader~\etal do not provide data for more than 50\% of the tasks.
    For two more, around 40\% of tasks have no historical runtime data stored.
    Hence, in the absence of historical data about a task, we give it fixed weights.
    We give it an execution time of $1$, a memory requirement of $50 MB$, and files written and received of $1KB$.
    These values align with the findings of~\cite{lotaru} about small tasks.
%
 %   \skug{insert more abt how the runtime system hanfles data reads}

    \subsubsection{Target computing systems}

    To fully benefit from the historical data, the  {\em default} experimental environment
    that we consider is a cluster based on the same six
    kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
    We set the number of each kind of node to $6$, thus having 36 processors in total.  % the whole cluster.
%
    Each machine has a normalized CPU speed and a normalized memory size, and we list them as (name, speed, memory):
    ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines;
    ($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and
    large memory (see Table~\ref{tab:procs}).

    \begin{table}[htb]
        \begin{center}
            \begin{tabular}{c|c|cc}
                \toprule
                Processor  %& Amount
                &  CPU speed   & \multicolumn{2}{c}{Memory size (relative)} \\
                name & (relative) & {\em default} & {\em mem-constrained} \\
                \midrule
                local                    & 4                    & 16     & 1.6 \\
                A1                      & 32                   & 32     & 3.2 \\
                A2                      & 6                    & 64    & 6.4 \\
                N1                      & 12                   & 16     & 1.6 \\
                N2                      & 8                    & 8      & 0.8\\
                C2                      & 32                   & 192   &  19.2\\
                \bottomrule
            \end{tabular}
        \end{center}
        \caption{Cluster configuration.}
        \label{tab:procs}
    \end{table}

    When we consider a default execution environment, we multiply the relative memory size with $10^6$.
    We multiply the relative speeds with $100$, and let read-write operations unpenalized (multiple the size of files with 1).
    Effectively, we thus reward the computations with $100$ in comparison  with IO operations.
    We further penalize offloading of memory by multiplying its execution time by $1000$.

    For a memory-constrained scenario, we give the processors ten time less memory: we multiply the relative memory size
    with $10^5$.
    We keep all the settings related to speeds the same.
    For a more memory-loo scenario, we give the processors ten times more memory.

    For the IO-based scenarios, we keep the amount of memory the same as in the default scenario and change only the penalties for the read-write
    operations and offloading.
    For the ``IO slower'' case, we penalize both the read-write operations and the offloading ten times more: we multiply the length of
    read-write operations by $10$, and the offloading by $10^4$.
    In the ``IO faster'' case, we do the opposite and reward read-write operations by dividing their time by to, while penalize the offloading only
    by $100$ instead of $1000$.
    Finally, we examine the effect of the offloading penalty reducing and increasing only it ten times for the experimental setups ``offload faster''
    and ``offload slower''.

    TODO: describe with more and less nodes.
   % We also consider a more constrained setting, by varying the cluster configuration. The
   %  {\em memory-constrained cluster} is consisting of 72 nodes (12 of each kind) as the default cluster,
   %  but each node has 10 times less memory. Hence,
   %  the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB,
   %%  $A2$ has $6.4$~GB instead, of $64$~GB etc.
   % The processor speeds and their relations stay unchanged (see Table~\ref{tab:procs}).

    \subsubsection{Runtime system}

    No ``runtime system'' as such in the static case.
    Todo: either remove entirely or explain how simulation of the runtime system works in the dynamic case.

\subsection{Results in a static setting} % \skug{Full: 3, Polished: 2}}
\label{sec.expe.static}

    We first study the heuristics behavior when the weights do not change upon runtime,
    hence the scheduler has a perfect knowledge of task memory requirements and execution times.
    We have compared two eviction strategies, starting with large files first or small files first,
    and did not report any significant changes in terms of validity of schedule or makespan.
    Hence, we only present results with the eviction of largest files first.

    \subsubsection{Relative makespan}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=0.9\columnwidth] {images/disk-model/rel-ms-normal-bar}
        \caption{Relative makespans on the default cluster setup. Lower is better.}
        \label{fig:rel-ms-normal-bare}
        \vspace{-0.3cm}
    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=0.9\columnwidth] {images/disk-model/rel-ms-iofast-bar}
        \caption{Relative makespans on the cluster setup with faster IO. Lower is better.}
        \label{fig:rel-ms-iofast-bar}
        \vspace{-0.3cm}
    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=0.9\columnwidth] {images/disk-model/rel-ms-ioslow-bar}
        \caption{Relative makespans on the cluster setup with slower IO. Lower is better.}
        \label{fig:rel-ms-ioslow-bar}
        \vspace{-0.3cm}
    \end{figure}

    \subsubsection{Scheduling on the default cluster}
    AFTER HERE IS OLD.
%    \paragraph{Normal cluster}

    On the default cluster,  the three memory-aware heuristics are able to schedule all workflows
    (see Figure~\ref{fig:success-rates-large}), while
    the baseline \heft has $24.2\%$ success rate ($75.7\%$ failure rate).
    Indeed, \heft is only able to schedule small workflows, but no workflow over the size of $4000$ tasks
    can be scheduled correctly; some tasks run out of memory. As soon as we are not in a setting
    with abundant processing resources for small workflows, it is hence necessary to adopt
    a memory-aware strategy in order to produce valid schedules.
%    Small workflows are less constrained by the memory requirement, so in the situation of abundant processing resources, even
%    a not memory-aware approach is able to produce valid schedules.
%    However, this advantage quickly vanishes as workflows grow larger.

    \begin{figure}[tb]
        \centering
        \includegraphics[width=0.9\columnwidth] {images/success-rates-large}
        \caption{Success rates on the default cluster. Higher is better.}
        \label{fig:success-rates-large}
        \vspace{-0.3cm}
    \end{figure}

   We also report in  Figure~\ref{fig:ms-relations-by-workflow}  the relative makespan found
   by the memory-aware heuristics, normalized to the
   makespan achieved by \heft, often through an invalid over-optimistic schedule that exceeds
   the bound on memory.
%    Figure~\ref{fig:ms-relations-by-workflow} shows the relations of the makespans found by the heuristics to the
%    makespans found by the baseline (including invalid ones), by workflow size.
    On average, the makespans found by \heftbl are $7.8\%$ worse than those found by the baseline,
    the makespans of \heftblc are $8\%$
    worse, and those found by \heftmm are $82.6\%$ worse. These are still very encouraging results,
    in particular for \heftbl and \heftblc, since the makespans of \heft correspond to invalid schedules.

 %   \skug{TODO: how much worse are makespans for SUCCESSFUL hefts? presulmably almost identical}

%    \begin{figure}[tb]
%        \centering
%        \includegraphics[width=1.1\columnwidth] {images/MsRelations2}
%        \caption{Relative makespans produced by the heuristics to the makespan produced by the baseline. Smaller is better.
%        \skug{SUGESTION: remove this picture, as it duplicates Fig 3}
%        }
%
%        \label{fig:ms-relations}
%        \vspace{-0.3cm}
%    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/makespan_relations_by_wf_size}
        \caption{Relative makespans of heuristics normalized by \heft makespan, by workflow
        size, on default cluster. Smaller is better.}
        \label{fig:ms-relations-by-workflow}
        \vspace{-0.3cm}
    \end{figure}

%    \paragraph{Memory Usage}

    Finally, we study the percentage of memory occupied by the schedule, which is another good indicator
    of the memory usage of the heuristic and its ability to produce valid schedules.  %its important characteristic.
    Figures \ref{fig:mem-usages-normal} and~\ref{fig:mem-usages-onlyvalid} show the percentage of memory
    occupied on average by the schedule
    produced by the different heuristics, for different workflow sizes, first on all schedules (hence including
    \heft schedules that were not valid), and then only on valid schedules (hence, no results for \heft
    on large workflow sizes).
%
%   Figure shows memory usages if only valid heft schedules are presented.
%    There are no valid heft scheduler from middle-sized and large workflows, so no red bars are present.

    \heftmm continuously outperforms other heuristics in terms of memory usage,
    using from $46\%$ less memory on the smallest
    workflows to $4$ times less memory on the largest \numprint{30000}-task workflows.
    If we consider the invalid \heft schedules, too, we see that they would require more and more memory on average,
    which explains why these schedules rapidly become invalid.
    This is because some assignments require more that $100\%$ of memory (which makes them invalid).
    We can assess the degree of invalidity by comparing \heft memory usage with the memory usage of \heftbl.
    \heftbl differs from the baseline only in the sense that it respects available memory on the processors.
    For the largest workflows, \heft schedules require almost twice the memory of \heftbl.

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal}
        \caption{Memory usage on default cluster, including invalid \heft schedules. }
 %by algorithm by workflow size
         \label{fig:mem-usages-normal}
        \vspace{-0.3cm}
    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/mem-usage-normal-onlyvalid}
        \caption{Memory usage on default cluster, including only valid \heft schedules. }
        \label{fig:mem-usages-onlyvalid}
        \vspace{-0.3cm}
    \end{figure}

\subsubsection{Scheduling on the memory-constrained cluster}

    On the memory-constrained cluster, \heft could produce valid assignments in only $14$ experiments
    out of $290$~($4.8\%$ success rate).
    The successful schedules were achieved exclusively on the tiny workflows (with only two 200-task size-increased
    workflow among them).
    \heftbl could successfully schedule $38\%$ of workflows, \heftblc could schedule $49\%$ of them, while \heftmm
    could still schedule all the workflows, including even the largest ones, see Figure~\ref{fig:success-rates-tiny}.
    As also observed on the normal cluster, \heftmm seems to be less affected by the size of the workflow
    when scheduling it than the other heuristics.

    Similarly to the default cluster, we observe that relative makespan of \heftmm is usually greater
    than those of \heft (see Fig.~\ref{fig:ms-relations-by-workflow-constrained}), but \heft schedules are
    almost all invalid. It is therefore very interesting to resort to \heftmm for large workflows
    in a constrained cluster, since tasks are processed in an order that minimizes
    the memory usage of schedule.
%
%    The makespans it produces are larger, but it seems to order the vertices in a similar way on both the small and the large workflows,
%    rather than resorting to more memory-consuming ordering that other heuristics and the baseline produces.
%    This makes it especially valuable for very large workflows in situations of very constrained memory in the cluster.
%
%

    Memory usages on the constrained cluster are depicted in~Fig.~\ref{fig:memory-usage-constrained},
    and we observe that the memory footprint of \heftmm remains constant with workflow size.

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1\columnwidth] {images/success-rates-tiny}
        \caption{Success rates on the memory-constrained cluster. Higher is better. }
%        by algorithm by workflow size.
        \label{fig:success-rates-tiny}
        \vspace{-0.3cm}
    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1\columnwidth] {images/makespan_relations_by_wf_size-constrained}
        \caption{Relative makespans on the memory-constrained cluster.
%        produced by the heuristics to the makespan produced by the baseline, by workflow size.
        Smaller is better.}
        \label{fig:ms-relations-by-workflow-constrained}
        \vspace{-0.3cm}
    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/mem-usage-constrained-onlyvalid}
        \caption{Memory usage on the memory-constrained  cluster. } % by different heuristics by workflow size.}
        \label{fig:memory-usage-constrained}
        \vspace{-0.3cm}
    \end{figure}

\subsection{Dynamic experiments}
\label{sec.expe.dyn}

    \paragraph{Memory-constrained cluster}
    The makespan in case of no recomputation becomes invalid as soon as at least one task finds itself in an invalid memory size
    situation - that is, if the scheduler assumed its memory to be smaller than it actually was and assigned it to a processor
    with not enough memory capacity.
    Due to an extremely constrained memory in this cluster, only 134 experiments out of 1160 for all algorithm variants
    ended with a valid makespan without recomputation.
    In case of \heft, $14$ valid initial makespans were computed.
    Out of them, $13$ managed to stay valid until the end, after all the update requests.
    The same $13$ experiments ended with a valid makespan in case of no recomputation.
    So, these workflows required so few resources that there was no point in re-scheduling them.
    %$13$ experiments ended with a valid makespan after recomputation in case of \heft, which corresponds to $4.4\%$ success
    %rate. However, there were only $14$ successful \heft experiments overall in this setup, so for the successful experiments
    %only, the success rate of not recomputing is $92\%$.
    In case of \heftmm, all $290$ workflows could be scheduled initially and all the schedules remained valid until the end.
    $16$ experiments ended with a successful makespan without recomputation.
    The rate of successful schedules without recomputation is therefore~$5.5\%$.
    \heftblc could produce $142$ valid initial schedules and kept $141$ of them valid until the end.
    $50$ experiments were successful without recomputation, too, making it $35\%$ out of all successful final schedules.
    \heftbl kept $105$ schedules valid until the end out of $110$ initial valid schedules.
    $55$ were able to be kept valid until the end even without recomputation, roughly $50\%$ of all successes.

    \heftbl and \heftblc were successful on smaller workflows and failed on larger ones, the success of the strategy without
    recomputation was limited to even the smallest of these smaller workflows.
    So, the strategy without recomputation delivered valid makespans (independently of the algorithm) on $56$ original
    workflows with $<100$ tasks, $47$ 200-task ones, $25$ 1000-task ones and $6$ 2000-task ones.

    Figure~\ref{fig:updates-ms} shows the increase in makespan in case of no recomputation for these experiments.
    With growing size of the workflow, the excess makespan of not recomputing grows - from $13.9\%$ to $20\%$ for \heftbl,
    from $12.7\%$ to $18.7\%$ on \heftblc (but there is no data for the 2000-task workflows in this case),
    $12.1\%$ to $23.5\%$ for \heftmm.
    The larger variations for \heft can be explained by the small amount of data -- for instance, there were only 2 200-task
    workflows for this case.

    \begin{figure}[tb]
        \centering
        \includegraphics[width=0.8\columnwidth] {images/UpdatesMss2}
        \caption{Relative (excess) makespan of \heftbl, \heftblc and \heftmm. Smaller is better.}
        \label{fig:updates-ms}
        \vspace{-0.3cm}
    \end{figure}

\subsection{Running times of the heuristics}
\label{sec.expe.t}

    To be able to answer the runtime system without holding it up for too long, the scheduler needs to 
    be able to compute a schedule rapidly. %provide a fast answer.
    The bottom-level-based heuristics \heftbl and \heftblc provide smaller running times than \heftmm, 
    and also scale better with growing workflow
    sizes~(see Fig.~\ref{fig:runtimes-log}).
    Their running times are similar and grow from tens of milliseconds for the smallest workflows to $25-27$ seconds 
    for the largest workflows.
    \heftmm, however, needs to compute a memory-optimal traversal of the entire workflow to compute the ranks 
    of the tasks,
    so its running time scales from also tens of milliseconds for the smallest workflows, 
    to thousands of seconds - $1172.7$ for \numprint{20000}-task workflows and
    $2994.9$ seconds for the largest, \numprint{30000}-task workflows.
    This increased running time is, though, offset by the unique $100\%$ success rate this algorithm 
    obtained when scheduling large
    workflows in difficult (memory-constrained) setups.

    \begin{figure}[tb]
        \centering
        \includegraphics[width=1.1\columnwidth] {images/runtimes-logarithmic}
        \caption{Running times of the heuristics. The y axis is logarithmic.}
        \label{fig:runtimes-log}
        \vspace{-0.3cm}
    \end{figure}

    %\begin{figure}[tb]
    %    \centering
    %    \includegraphics[width=1.1\columnwidth] {images/runtimes-absolute}
    %    \caption{Running times of the heuristics and the baseline.}
    %    \label{fig:runtimes-abs}
    %    \vspace{-0.3cm}
    %\end{figure}

%    \subsubsection{Summary}

\section{Conclusion} %  \skug{Full: 0, Polished: 0}}
\label{sec:conc}

We have formalized a scheduling problem in memory-constrained environments, where tasks
may exceed the memory available on a processor and resort to communication buffers
to store and communicate data between processors. In this context, we have 
designed three memory-aware HEFT-based heuristics, which account for memory
constraints while scheduling tasks.

Two heuristics rely on an ordering of tasks based on the bottom level of tasks, with the objective
of minimizing the makespan and hence scheduling critical tasks first. The third one, \heftmm,
goes one step further and handles tasks in an order dictated by an efficient traversal
of the graph in terms of memory requirements, hence reducing the memory used
by the schedule. Experimental results on a large set of workflows coming from real-life applications
demonstrate that the memory-aware heuristics are  successfully producing valid schedules. 
In the most memory-constrained setting, \heftmm succeeds to schedule even the largest workflows,
while other heuristics return invalid schedules exceeding the memory capacity of processors. 
It comes at a price of a makespan that is not quite as good as the one obtained by \heftbl
and \heftblc, which focus on makespan by ordering tasks by bottom level. 
As expected, the baseline \heft that is not memory aware returns invalid schedules in almost
all settings, except for very small workflows.

%Hence, \heftbl and \heftblc can schedule all workflows in a normal cluster and approximately half of them in a memory-constrained
%    one.
%    They provide a running time that scales well with the growing size of a workflow.
%
%    \heftmm can schedule even the largest workflows even on the memory-constrained cluster.
%    It is the only heuristic that has $100\%$ success rate in our experiments.
%    However, the makespans it produces are larger than those of the the other two.

Another key contribution is that we have adapted these heuristics for a dynamic setting, where exact
task parameters (execution times and memory requirements) are not know in advance. 
We have implemented a runtime system that interacts with the scheduler, returning exact
parameter values once a task arrives in the system, while only estimates are known for future tasks. 
Some preliminary experiments have been conducted in this setting, and demonstrated that
it is necessary to adapt the schedule on the fly in order to avoid an execution failure because
of a shortage in memory. To the best of our knowledge, this is the first study of adaptive
algorithms accounting for memory constraints.

This work could be extended in several directions. First, the model could be refined to include
heterogeneous bandwidths, while we consider a homogeneous communication network. 
More importantly, it would be interesting to consider other types of variability, for instance
if new tasks appear in the graph (or disappear), or if there is variability in the platform, 
with processors arriving and departing. We believe that we could adapt the current approach
by recomputing schedules on the fly, and we plan to perform a new set of experiments
to further assess the impact of dynamic scheduling.

%
%\begin{itemize}
%  \item Extend to heterogeneous bandwidths. 
%  \item Further variability commented out from the paper: changing DAG structure; changing platform...
%\end{itemize}

\balance
    \bibliographystyle{abbrv}%IEEEtran}
    \bibliography{references}

\end{document}